---
title: 'TRAP Prediction Model Part \MakeUppercase{\romannumeral 2}: Prediction at Participant Locations'
author: "Si Cheng"
date: ' `r Sys.Date()` '
geometry: margin=.75in
output:
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
load("pred_cohort.RData")
```

## Prediction maps of each method
We run each of the models obtained from training data to make predictions at the participant locations, visualized in the plots below. We focus on the unweighted annual averages of UFP and BC concentration.

```{r, fig.width=15, fig.height=12}
library(sf)
library(tidyverse)
library(ggmap)
library(grid)
library(ggplot2)
library(gridExtra)

project_crs <- 4326

# cohort_covars_shp <- cov_cohort %>%
#   st_as_sf(., coords=c("longitude","latitude"), remove = FALSE,
#            crs=project_crs)

seaMap <- get_map(location=c(min(cov_cohort$longitude), min(cov_cohort$latitude),
                            max(cov_cohort$longitude), max(cov_cohort$latitude)),
                 source="stamen", maptype = "terrain", crop=FALSE)

plot_pred = function(mod, title){
  mod$Y.pls = mod$Y.pls
  mod$Y.spatrf.pl = mod$Y.spatrf.pl 
  mod$Y.spatrf.cv = mod$Y.spatrf.cv 
  mod$Y.tprs = mod$Y.tprs 
  mod$Y.tprs.rf = mod$Y.tprs.rf 
  mod$Y.rf = mod$Y.rf 
  mod$Y.rf.tprs = mod$Y.rf.tprs 
  mod$Y.pls.mean = mod$Y.pls.mean 
  mod$Y.pls.cov = mod$Y.pls - mod$Y.pls.mean
  
  plot_range = range(c(mod$Y.pls, mod$Y.spatrf.pl, mod$Y.spatrf.cv,
                       mod$Y.tprs, mod$Y.tprs.rf, mod$Y.rf, mod$Y.rf.tprs,
                       mod$Y.pls.cov))#,
                       #mod$Y.pls.mean))
  temp = data.frame(longitude.x = cov_cohort$longitude , 
                    latitude.x = cov_cohort$latitude ,
                    Y.pls = mod$Y.pls, Y.spatrf.pl = mod$Y.spatrf.pl,
                    Y.spatrf.cv = mod$Y.spatrf.cv, Y.tprs = mod$Y.tprs,
                    Y.tprs.rf = mod$Y.tprs.rf, Y.rf = mod$Y.rf, 
                    Y.rf.tprs = mod$Y.rf.tprs, 
                    Y.pls.mean = mod$Y.pls.mean,
                    Y.pls.cov = mod$Y.pls.cov)
  
  p_pls = ggmap(seaMap) + labs(x = "longitude", y = "latitude") +
            geom_point(data = temp, aes(x = longitude.x, y = latitude.x, 
                             color = Y.pls), shape = ".") + #size = Y.pls
            #scale_color_viridis_c(option = "plasma", name = "", 
            #                      limits = plot_range, alpha = .8) +
            scale_color_gradient(low = "white", high = "red",
                                 name = "", limits = plot_range) +
            ggtitle("UK PLS") + guides(size = FALSE) +
            theme(plot.margin=unit(rep(0,4), "pt"), 
                  text = element_text(size = 16))
  p_pls.mean = ggmap(seaMap) + labs(x = "longitude", y = "latitude") +
            geom_point(data = temp, aes(x = longitude.x, y = latitude.x, 
                             color = Y.pls.mean), shape = ".") + #size = Y.pls
            #scale_color_viridis_c(option = "plasma", name = "", 
            #                      limits = plot_range, alpha = .8) +
            scale_color_gradient(low = "white", high = "red", name = "") +
                                 #name = "", limits = plot_range) +
            ggtitle("UK PLS (Mean)") + guides(size = FALSE) +
            theme(plot.margin=unit(rep(0,4), "pt"), 
                  text = element_text(size = 16))
  p_pls.cov = ggmap(seaMap) + labs(x = "longitude", y = "latitude") +
            geom_point(data = temp, aes(x = longitude.x, y = latitude.x, 
                             color = Y.pls.cov), shape = ".") + #size = Y.pls
            #scale_color_viridis_c(option = "plasma", name = "", 
            #                      limits = plot_range, alpha = .8) +
            scale_color_gradient(low = "white", high = "red",
                                 name = "", limits = plot_range) +
            ggtitle("UK PLS (Cov)") + guides(size = FALSE) +
            theme(plot.margin=unit(rep(0,4), "pt"), 
                  text = element_text(size = 16))
  p_spatrf.pl = ggmap(seaMap) + labs(x = "longitude", y = "latitude") +
            geom_point(data = temp, aes(x = longitude.x, y = latitude.x, 
                             color = Y.spatrf.pl), shape = ".") + #size = Y.pls
            scale_color_gradient(low = "white", high = "red",
                                 name = "", limits = plot_range) +
            #geom_point(data = datsub, aes(x = longitude.x, y = latitude.x, 
            #                 color = res.pls, size = abs(res.pls))) +
            #scale_color_gradient2(name = "", limits = c(lower, upper)) +
            ggtitle("SpatRF-PL") + guides(size = FALSE) +
            theme(plot.margin=unit(rep(0,4), "pt"), 
                  text = element_text(size = 16))
  p_spatrf.cv = ggmap(seaMap) + labs(x = "longitude", y = "latitude") +
            geom_point(data = temp, aes(x = longitude.x, y = latitude.x, 
                             color = Y.spatrf.cv), shape = ".") + #size = Y.pls
            scale_color_gradient(low = "white", high = "red",
                                 name = "", limits = plot_range) +
            #geom_point(data = datsub, aes(x = longitude.x, y = latitude.x, 
            #                 color = res.pls, size = abs(res.pls))) +
            #scale_color_gradient2(name = "", limits = c(lower, upper)) +
            ggtitle("SpatRF-CV") + guides(size = FALSE) +
            theme(plot.margin=unit(rep(0,4), "pt"), 
                  text = element_text(size = 16))
  p_tprs = ggmap(seaMap) + labs(x = "longitude", y = "latitude") +
            geom_point(data = temp, aes(x = longitude.x, y = latitude.x, 
                             color = Y.tprs), shape = ".") + #size = Y.pls
            scale_color_gradient(low = "white", high = "red",
                                 name = "", limits = plot_range) +
            #geom_point(data = datsub, aes(x = longitude.x, y = latitude.x, 
            #                 color = res.pls, size = abs(res.pls))) +
            #scale_color_gradient2(name = "", limits = c(lower, upper)) +
            ggtitle("TPRS") + guides(size = FALSE) +
            theme(plot.margin=unit(rep(0,4), "pt"), 
                  text = element_text(size = 16))
  p_rf = ggmap(seaMap) + labs(x = "longitude", y = "latitude") +
            geom_point(data = temp, aes(x = longitude.x, y = latitude.x, 
                             color = Y.rf), shape = ".") + #size = Y.pls
            scale_color_gradient(low = "white", high = "red",
                                 name = "", limits = plot_range) +
            ggtitle("RF") + guides(size = FALSE) +
            theme(plot.margin=unit(rep(0,4), "pt"), 
                  text = element_text(size = 16))
  p_tprs.rf = ggmap(seaMap) + labs(x = "longitude", y = "latitude") +
            geom_point(data = temp, aes(x = longitude.x, y = latitude.x, 
                             color = Y.tprs.rf), shape = ".") + #size = Y.pls
            scale_color_gradient(low = "white", high = "red",
                                 name = "", limits = plot_range) +
            ggtitle("TPRS-RF") + guides(size = FALSE) +
            theme(plot.margin=unit(rep(0,4), "pt"), 
                  text = element_text(size = 16))
  p_rf.tprs = ggmap(seaMap) + labs(x = "longitude", y = "latitude") +
            geom_point(data = temp, aes(x = longitude.x, y = latitude.x, 
                             color = Y.rf.tprs), shape = ".") + #size = Y.pls
            scale_color_gradient(low = "white", high = "red",
                                 name = "", limits = plot_range) +
            ggtitle("RF-TPRS") + guides(size = FALSE) +
            theme(plot.margin=unit(rep(0,4), "pt"), 
                  text = element_text(size = 16))
  
  grid.arrange(arrangeGrob(p_rf, p_tprs, p_tprs.rf, p_rf.tprs,
                           p_spatrf.pl, p_spatrf.cv, 
                           p_pls, p_pls.mean, p_pls.cov,
                           nrow=3),
               top = textGrob(title,gp=gpar(fontsize=30)))
}

plot_pred(grid.ufp.uw, "UFP (unweighted)")
```

```{r, fig.width=15, fig.height=12}
plot_pred(grid.bc.uw, "BC (unweighted)")
```


## Variable importance of PLS and SpatRF
As in Part \MakeUppercase{\romannumeral 1}, we visualize the importance of each predictor in PLS, UK-PLS, random forest and spatial random forest models, defined as

* PLS and UK-PLS: linear combination of loadings based on PLS regression parameters
* RF and SpatRF: proportion/amount of loss accounted for by making splits based on each variable (SpatRF normalizes by the maximum loss, while RF does not)

```{r, fig.height=10, fig.width=15}
dat = readRDS("dat.rda")
grid <- list(lambert_x = dat$lambert_x,
             lambert_y = dat$lambert_y)
X = dat[,-(1:21)]
source("C:/Users/Si Cheng/Dropbox (Personal)/exposurepred/spatialRF/aux_functions.R")
source("spatTreeModified.R")

getImp = function(rslt, y){
  pls.loadings = rslt$pc.obj$loadings
  new.x.train <- rslt$pc.obj$X
  theta.pls <- rslt$theta.pls
  sig.pls <- exp(theta.pls[1])
  nug.pls <- exp(theta.pls[2])
  range.pls <- exp(theta.pls[3])
  dmat <- as.matrix(SpatioTemporal::crossDist(cbind(grid$lambert_x,
                                                    grid$lambert_y)))
  i.V <- .fastSolve(sig.pls * exp(-dmat/range.pls) + 
                      nug.pls * diag(length(y)))
  beta.hat.pcr <- solve(t(new.x.train) %*% i.V %*% new.x.train,
                        t(new.x.train) %*% i.V %*% y)
  beta.hat.indep <- solve(t(new.x.train) %*% new.x.train,
                        t(new.x.train) %*% y)
  imp.pls = pls.loadings %*% beta.hat.pcr
  # imp.pls.indep = coef(pls::plsr(y ~ as.matrix(X), 
  #                 ncomp = ncol(rslt$pc.obj$loadings),scale=TRUE),
  #                 ncomp = ncol(rslt$pc.obj$loadings))
  imp.pls.indep = pls.loadings %*% beta.hat.indep
  imp.spatrf.cv = rslt$spatrf.cv$var.imp
  imp.spatrf.pl = rslt$spatrf.pl$var.imp
  imp.rf = rslt$rf.imp[,2]
  return(imp = cbind(imp.pls, imp.pls.indep, imp.spatrf.cv, 
                     imp.spatrf.pl, imp.rf))
}

# imp.ufp.primary = getImp(grid.ufp.primary, dat$ufp_primary)
imp.ufp.uw = getImp(grid.ufp.uw, dat$ufp_uw)
# imp.bc.primary = getImp(grid.bc.primary, dat$bc_primary)
imp.bc.uw = getImp(grid.bc.uw, dat$bc_uw)

plotImp = function(impMat, tt){
  my.alpha = .3
  colnames(impMat) = c("Method PLS", "Method PLS-Indep", 
                       "Method SpatRF-CV", "Method SpatRF-PL", "Method RF")
  
  split_cov_name <- function(dt, cov) {
    dt <- suppressWarnings(dt %>%
                             rename(cov_full_name = cov) %>%
                             mutate(
                               buffer = substr(cov_full_name, nchar(cov_full_name)-4, nchar(cov_full_name)),
                               buffer = as.numeric(ifelse(!is.na(as.integer(buffer)), buffer, NA)),
                               cov = ifelse(is.na(buffer), cov_full_name, substr(cov_full_name, 1, nchar(cov_full_name)-5) )
                             )
    ) %>%
      select(contains("cov"), buffer, everything())
    
    # elevation
    dt$cov[grepl("^elev_.+_above$", dt$cov_full_name)] <- "elev_above"
    dt$cov[grepl("^elev_.+_below$", dt$cov_full_name)] <- "elev_below"
    dt$cov[grepl("^elev_.+_stdev$", dt$cov_full_name)] <- "elev_stdev"
    dt$cov[grepl("^elev_.+_at_elev$", dt$cov_full_name)] <- "elev_at_elev"
    
    dt$buffer[grepl("_1k_", dt$cov_full_name)] <- 1000
    dt$buffer[grepl("_5k_", dt$cov_full_name)] <- 5000
    
    return(dt)
  }
  
  impMat = as.matrix(impMat)
  rownames(impMat) = colnames(X)
  dtNew <- impMat %>%
    as.data.frame() %>%
    rownames_to_column(var = "cov") %>%
    # rename variables if buffers
    split_cov_name(cov = "cov") %>%
      #make long format for faceting
      gather(key = "Method", value = "Importance", contains("Method")) %>%
      mutate(Method = substr(Method, 8, nchar(Method))) 
  #colnames(dtNew)[4] = "Importance"
  
  dtNew  %>%
    #buffered covariates
    drop_na(buffer) %>%
    ggplot(aes(x = Importance, y = cov)) +
    geom_point(aes(size=buffer),
               shape=1,
               alpha=my.alpha) +
    scale_size(breaks = c(min(dtNew$buffer, na.rm = T),
                          max(dtNew$buffer, na.rm = T)
    )) + #500, 5000, 10000,
    #non-buffered covariates
    geom_point(data = dtNew[is.na(dtNew$buffer),],
               alpha=my.alpha,
               aes(shape="")) +
    geom_vline(xintercept=0,
               linetype="solid",
               alpha=my.alpha) +
    facet_wrap(~ Method, scales = "free_x",
                 labeller = "label_both",
                 nrow = 1) +
    labs(y = "Geocovariate",
         shape= "non-buffer", #"proximity,\nelevation",
         #title = paste("PLS Importance:", pollutant)) +
         title = tt) +
    theme(legend.position = "bottom", text = element_text(size = 16))
}


# plotImp(imp.ufp.primary, "UFP (adjusted)")
plotImp(imp.ufp.uw, "UFP (unweighted)")
# plotImp(imp.bc.primary, "BC (adjusted)")
plotImp(imp.bc.uw, "BC (unweighted)")
```

Pairwise scatter plots compare the (relative) magnitude of variable importance in different prediction models. Absolute values were taken for PLS importance measures.

```{r}
# pairs(abs(imp.ufp.primary), pch = 19, main = "UFP (adjusted)",
#       labels = c("PLS", "PLS-indep", "SpatRF-CV", "SpatRF-PL", "RF"))
pairs(abs(imp.ufp.uw), pch = 19, main = "UFP (unweighted)",
      labels = c("PLS", "PLS-indep", "SpatRF-CV", "SpatRF-PL", "RF"))
# pairs(abs(imp.bc.primary), pch = 19, main = "BC (adjusted)",
#       labels = c("PLS", "PLS-indep", "SpatRF-CV", "SpatRF-PL", "RF"))
pairs(abs(imp.bc.uw), pch = 19, main = "BC (unweighted)",
      labels = c("PLS", "PLS-indep", "SpatRF-CV", "SpatRF-PL", "RF"))
```
