---
title: "TRAP model: grid predictions and variable importance plots"
author: "Si Cheng"
date: ' `r Sys.Date()` '
output: 
  html_document:
    number_sections: no
    df_print: paged
    toc: yes
    toc_depth: '3'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
load("rslt_grid_uw_mean.RData")
grid.bc.uw.mean = grid.bc.uw
grid.ufp.uw.mean = grid.ufp.uw
load("rslt_grid_uw2.RData")
pls.bc.uw.mean = grid.bc.uw
pls.ufp.uw.mean = grid.ufp.uw
load("rslt_grid_uw.RData")
grid.bc.uw$Y.pls.mean = pls.bc.uw.mean$Y.pls.mean
grid.ufp.uw$Y.pls.mean = pls.ufp.uw.mean$Y.pls.mean
grid.bc.uw$Y.pls.cov = pls.bc.uw.mean$Y.pls - pls.bc.uw.mean$Y.pls.mean
grid.ufp.uw$Y.pls.cov = pls.ufp.uw.mean$Y.pls - pls.ufp.uw.mean$Y.pls.mean
```

## Grid predictions
```{r, fig.width=16, fig.height=12}
library(sf)
library(tidyverse)
library(ggmap)
library(grid)
library(ggplot2)
library(gridExtra)

project_crs <- 4326
study_area_shp <- read_sf("oval_shp/oval_around_monitoring_area.shp") %>% 
  st_transform(project_crs)

grid_covars_shp <- cov_grid %>%
  st_as_sf(., coords=c("longitude","latitude"), remove = FALSE,
           crs=project_crs)
grid_covars_shp$in_study_area <- st_intersects(grid_covars_shp, study_area_shp,
                                               sparse = FALSE) %>%
  apply(., 1, any)


seaMap <- get_map(location=c(min(cov_grid$longitude), min(cov_grid$latitude),
                            max(cov_grid$longitude), max(cov_grid$latitude)),
                 source="stamen", maptype = "terrain", crop=FALSE)
#seaMap2 <- get_map(location=c(min(cov_grid$longitude), min(cov_grid$latitude),
#                            max(cov_grid$longitude), max(cov_grid$latitude)),
#                   source = "stamen", maptype = "watercolor", crop=FALSE)

plot_pred = function(mod, mod.mean, title){
  mod$Y.pls = mod$Y.pls[grid_covars_shp$in_study_area]
  mod$Y.spatrf.pl = mod$Y.spatrf.pl[grid_covars_shp$in_study_area]
  mod$Y.spatrf.cv = mod$Y.spatrf.cv[grid_covars_shp$in_study_area]
  mod$Y.tprs = mod$Y.tprs[grid_covars_shp$in_study_area]
  mod$Y.tprs.rf = mod$Y.tprs.rf[grid_covars_shp$in_study_area]
  mod$Y.rf = mod$Y.rf[grid_covars_shp$in_study_area]
  mod$Y.rf.tprs = mod$Y.rf.tprs[grid_covars_shp$in_study_area]
  mod$Y.pls.mean = mod$Y.pls.mean[grid_covars_shp$in_study_area]
  mod$Y.pls.cov = mod$Y.pls.cov[grid_covars_shp$in_study_area]
  mod.mean$Y.spatrf.pl = mod.mean$Y.spatrf.pl[grid_covars_shp$in_study_area]
  mod.mean$Y.spatrf.cv = mod.mean$Y.spatrf.cv[grid_covars_shp$in_study_area]
  
  plot_range = range(c(mod$Y.pls, mod$Y.spatrf.pl, mod$Y.spatrf.cv,
                       mod$Y.tprs, mod$Y.tprs.rf, mod$Y.rf, mod$Y.rf.tprs,
                       mod.mean$Y.spatrf.pl, mod.mean$Y.spatrf.cv,
                       mod$Y.pls.cov))#,
                       #mod$Y.pls.mean))
  temp = data.frame(longitude.x = cov_grid$longitude[grid_covars_shp$in_study_area], 
                    latitude.x = cov_grid$latitude[grid_covars_shp$in_study_area],
                    Y.pls = mod$Y.pls, Y.spatrf.pl = mod$Y.spatrf.pl,
                    Y.spatrf.cv = mod$Y.spatrf.cv, Y.tprs = mod$Y.tprs,
                    Y.tprs.rf = mod$Y.tprs.rf, Y.rf = mod$Y.rf, 
                    Y.rf.tprs = mod$Y.rf.tprs, 
                    Y.spatrf.cv.mean = mod.mean$Y.spatrf.cv,
                    Y.spatrf.pl.mean = mod.mean$Y.spatrf.pl,
                    Y.pls.mean = mod$Y.pls.mean,
                    Y.pls.cov = mod$Y.pls.cov)
  
  p_pls = ggmap(seaMap) + labs(x = "longitude", y = "latitude") +
            geom_point(data = temp, aes(x = longitude.x, y = latitude.x, 
                             color = Y.pls)) + #size = Y.pls
            #scale_color_viridis_c(option = "plasma", name = "", 
            #                      limits = plot_range, alpha = .8) +
            scale_color_gradient(low = "white", high = "red",
                                 name = "", limits = plot_range) +
            ggtitle("UK PLS") + guides(size = FALSE) +
            theme(plot.margin=unit(rep(0,4), "pt"), 
                  text = element_text(size = 16))
  p_pls.mean = ggmap(seaMap) + labs(x = "longitude", y = "latitude") +
            geom_point(data = temp, aes(x = longitude.x, y = latitude.x, 
                             color = Y.pls.mean)) + #size = Y.pls
            #scale_color_viridis_c(option = "plasma", name = "", 
            #                      limits = plot_range, alpha = .8) +
            scale_color_gradient(low = "white", high = "red", name = "") +
                                 #name = "", limits = plot_range) +
            ggtitle("UK PLS (Mean)") + guides(size = FALSE) +
            theme(plot.margin=unit(rep(0,4), "pt"), 
                  text = element_text(size = 16))
  p_pls.cov = ggmap(seaMap) + labs(x = "longitude", y = "latitude") +
            geom_point(data = temp, aes(x = longitude.x, y = latitude.x, 
                             color = Y.pls.cov)) + #size = Y.pls
            #scale_color_viridis_c(option = "plasma", name = "", 
            #                      limits = plot_range, alpha = .8) +
            scale_color_gradient(low = "white", high = "red",
                                 name = "", limits = plot_range) +
            ggtitle("UK PLS (Cov)") + guides(size = FALSE) +
            theme(plot.margin=unit(rep(0,4), "pt"), 
                  text = element_text(size = 16))
  p_spatrf.pl = ggmap(seaMap) + labs(x = "longitude", y = "latitude") +
            geom_point(data = temp, aes(x = longitude.x, y = latitude.x, 
                             color = Y.spatrf.pl)) + #size = Y.pls
            scale_color_gradient(low = "white", high = "red",
                                 name = "", limits = plot_range) +
            #geom_point(data = datsub, aes(x = longitude.x, y = latitude.x, 
            #                 color = res.pls, size = abs(res.pls))) +
            #scale_color_gradient2(name = "", limits = c(lower, upper)) +
            ggtitle("SpatRF-PL") + guides(size = FALSE) +
            theme(plot.margin=unit(rep(0,4), "pt"), 
                  text = element_text(size = 16))
  p_spatrf.cv = ggmap(seaMap) + labs(x = "longitude", y = "latitude") +
            geom_point(data = temp, aes(x = longitude.x, y = latitude.x, 
                             color = Y.spatrf.cv)) + #size = Y.pls
            scale_color_gradient(low = "white", high = "red",
                                 name = "", limits = plot_range) +
            #geom_point(data = datsub, aes(x = longitude.x, y = latitude.x, 
            #                 color = res.pls, size = abs(res.pls))) +
            #scale_color_gradient2(name = "", limits = c(lower, upper)) +
            ggtitle("SpatRF-CV") + guides(size = FALSE) +
            theme(plot.margin=unit(rep(0,4), "pt"), 
                  text = element_text(size = 16))
  p_tprs = ggmap(seaMap) + labs(x = "longitude", y = "latitude") +
            geom_point(data = temp, aes(x = longitude.x, y = latitude.x, 
                             color = Y.tprs)) + #size = Y.pls
            scale_color_gradient(low = "white", high = "red",
                                 name = "", limits = plot_range) +
            #geom_point(data = datsub, aes(x = longitude.x, y = latitude.x, 
            #                 color = res.pls, size = abs(res.pls))) +
            #scale_color_gradient2(name = "", limits = c(lower, upper)) +
            ggtitle("TPRS") + guides(size = FALSE) +
            theme(plot.margin=unit(rep(0,4), "pt"), 
                  text = element_text(size = 16))
  p_rf = ggmap(seaMap) + labs(x = "longitude", y = "latitude") +
            geom_point(data = temp, aes(x = longitude.x, y = latitude.x, 
                             color = Y.rf)) + #size = Y.pls
            scale_color_gradient(low = "white", high = "red",
                                 name = "", limits = plot_range) +
            ggtitle("RF") + guides(size = FALSE) +
            theme(plot.margin=unit(rep(0,4), "pt"), 
                  text = element_text(size = 16))
  p_tprs.rf = ggmap(seaMap) + labs(x = "longitude", y = "latitude") +
            geom_point(data = temp, aes(x = longitude.x, y = latitude.x, 
                             color = Y.tprs.rf)) + #size = Y.pls
            scale_color_gradient(low = "white", high = "red",
                                 name = "", limits = plot_range) +
            ggtitle("TPRS-RF") + guides(size = FALSE) +
            theme(plot.margin=unit(rep(0,4), "pt"), 
                  text = element_text(size = 16))
  p_rf.tprs = ggmap(seaMap) + labs(x = "longitude", y = "latitude") +
            geom_point(data = temp, aes(x = longitude.x, y = latitude.x, 
                             color = Y.rf.tprs)) + #size = Y.pls
            scale_color_gradient(low = "white", high = "red",
                                 name = "", limits = plot_range) +
            ggtitle("RF-TPRS") + guides(size = FALSE) +
            theme(plot.margin=unit(rep(0,4), "pt"), 
                  text = element_text(size = 16))
  
  p_spatrf.pl.mean = ggmap(seaMap) + labs(x = "longitude", y = "latitude") +
            geom_point(data = temp, aes(x = longitude.x, y = latitude.x, 
                             color = Y.spatrf.pl.mean)) + #size = Y.pls
            scale_color_gradient(low = "white", high = "red",
                                 name = "", limits = plot_range) +
            #geom_point(data = datsub, aes(x = longitude.x, y = latitude.x, 
            #                 color = res.pls, size = abs(res.pls))) +
            #scale_color_gradient2(name = "", limits = c(lower, upper)) +
            ggtitle("SpatRF-PL (Mean)") + guides(size = FALSE) +
            theme(plot.margin=unit(rep(0,4), "pt"), 
                  text = element_text(size = 16))
  p_spatrf.cv.mean = ggmap(seaMap) + labs(x = "longitude", y = "latitude") +
            geom_point(data = temp, aes(x = longitude.x, y = latitude.x, 
                             color = Y.spatrf.cv.mean)) + #size = Y.pls
            scale_color_gradient(low = "white", high = "red",
                                 name = "", limits = plot_range) +
            #geom_point(data = datsub, aes(x = longitude.x, y = latitude.x, 
            #                 color = res.pls, size = abs(res.pls))) +
            #scale_color_gradient2(name = "", limits = c(lower, upper)) +
            ggtitle("SpatRF-CV (Mean)") + guides(size = FALSE) +
            theme(plot.margin=unit(rep(0,4), "pt"), 
                  text = element_text(size = 16))
  grid.arrange(arrangeGrob(p_rf, p_tprs, p_tprs.rf, p_rf.tprs, 
                           p_spatrf.pl, p_spatrf.cv, 
                           p_spatrf.pl.mean, p_spatrf.cv.mean, 
                           p_pls, p_pls.mean, p_pls.cov,
                           nrow=3),
               top = textGrob(title,gp=gpar(fontsize=30)))
}
plot_pred(grid.ufp.uw, grid.ufp.uw.mean, "UFP (unweighted)")
plot_pred(grid.bc.uw, grid.bc.uw.mean, "BC (unweighted)")
```

### Pairwise scatterplot of predicted values

```{r}
# pairs(abs(imp.ufp.primary), pch = 19, main = "UFP (adjusted)",
#       labels = c("PLS", "PLS-indep", "SpatRF-CV", "SpatRF-PL", "RF"))
twolines = function(x,y) {
    points(x,y,pch=".")
    abline(a=0,b=1,col="red")
}

pairs(cbind(grid.ufp.uw$Y.pls[grid_covars_shp$in_study_area],
            pls.ufp.uw.mean$Y.pls.mean[grid_covars_shp$in_study_area],
            grid.ufp.uw$Y.spatrf.cv[grid_covars_shp$in_study_area],
            grid.ufp.uw$Y.spatrf.pl[grid_covars_shp$in_study_area],
            grid.ufp.uw$Y.rf[grid_covars_shp$in_study_area],
            grid.ufp.uw$Y.rf.tprs[grid_covars_shp$in_study_area]), 
      panel = twolines, main = "Predictions: UFP (unweighted)", 
      labels = c("PLS", "PLS (mean)", "SpatRF-CV", "SpatRF-PL", "RF", "RF-TPRS"))
# pairs(abs(imp.bc.primary), pch = 19, main = "BC (adjusted)",
#       labels = c("PLS", "PLS-indep", "SpatRF-CV", "SpatRF-PL", "RF"))
pairs(cbind(grid.bc.uw$Y.pls[grid_covars_shp$in_study_area],
            pls.bc.uw.mean$Y.pls.mean[grid_covars_shp$in_study_area],
            grid.bc.uw$Y.spatrf.cv[grid_covars_shp$in_study_area],
            grid.bc.uw$Y.spatrf.pl[grid_covars_shp$in_study_area],
            grid.bc.uw$Y.rf[grid_covars_shp$in_study_area],
            grid.bc.uw$Y.rf.tprs[grid_covars_shp$in_study_area]), 
      panel = twolines, main = "Predictions: BC (unweighted)", 
      labels = c("PLS", "PLS (mean)", "SpatRF-CV", "SpatRF-PL", "RF", "RF-TPRS"))
```

```{r, fig.height=12, fig.width=15}
## header: ## Plots of some covariates

# plot_covariate = function(name, title, takeLog = TRUE){
#   dat_temp = cov_grid
#   dat_temp$obs = cov_grid[,which(colnames(cov_grid) == name)]
#   if (takeLog) dat_temp$obs = log(dat_temp$obs + 1)
#   p_cov = ggmap(seaMap) + labs(x = "longitude", y = "latitude") +
#             geom_point(data = dat_temp, aes(x = longitude, y = latitude, 
#                              color = obs)) + #, size = obs
#             scale_color_viridis_c(option = "plasma", name = "") +# + upper)) + 
#             guides(size = FALSE) + ggtitle(title) +
#             theme(plot.margin=unit(rep(0,4), "pt"), 
#                   text = element_text(size = 16))
#   p_cov
# }
# pcov1 = plot_covariate("tl_s01000", "Truck routes (log), buffer = 1km", TRUE)
# pcov2 = plot_covariate("log_m_to_truck", "Distance to truck routes (log)", FALSE)
# pcov3 = plot_covariate("ll_a1_s05000", "A1 roads (log), buffer = 5km", TRUE)
# pcov4 = plot_covariate("log_m_to_a1", "Distance to A1 roads (log)", FALSE)
# pcov5 = plot_covariate("log_m_to_a1_a3_intersect", 
#                        "Distance to A1 A3 intersection (log)", FALSE)
# pcov6 = plot_covariate("elev_elevation", "Elevation", FALSE)
# grid.arrange(arrangeGrob(pcov1, pcov2, pcov3, pcov4, pcov5, pcov6, nrow=2))
```


## Variable importance plots
The plots below visualize the ``importance'' of each predictor, defined as

* PLS: linear combination of loadings based on PLS regression parameters
* SpatRF: proportion of loss accounted for by making splits based on each variable

```{r, fig.height=10, fig.width=15}
load("pred_grid_imp.RData")

getImp = function(rslt, y){
  pls.loadings = rslt$pc.obj$loadings
  new.x.train <- rslt$pc.obj$X
  theta.pls <- rslt$theta.pls
  sig.pls <- exp(theta.pls[1])
  nug.pls <- exp(theta.pls[2])
  range.pls <- exp(theta.pls[3])
  
  train.coords <- cbind(dat$lambert_x, dat$lambert_y)
  scale.params <- cbind(apply(train.coords,2,mean),apply(train.coords,2,sd))
  train.coords[,1] <- (train.coords[,1] - scale.params[1,1])/scale.params[1,2]
  train.coords[,2] <- (train.coords[,2] - scale.params[2,1])/scale.params[2,2]
  
  dmat <- as.matrix(SpatioTemporal::crossDist(train.coords))
  i.V <- .fastSolve(sig.pls * exp(-dmat/range.pls) + 
                      nug.pls * diag(length(y)))
  beta.hat.pcr = solve(t(new.x.train) %*% i.V %*% new.x.train,
                        t(new.x.train) %*% i.V %*% y)
  beta.hat.indep = solve(t(new.x.train) %*% new.x.train,
                        t(new.x.train) %*% y)
  imp.pls = rslt$pc.obj$proj %*% beta.hat.pcr
  #imp.pls = pls.loadings %*% beta.hat.pcr
  imp.pls.indep = coef(pls::plsr(y ~ as.matrix(X),
                  ncomp = ncol(rslt$pc.obj$loadings),scale=TRUE),
                  ncomp = ncol(rslt$pc.obj$loadings))
  # imp.pls.indep2 = rslt$pc.obj$proj %*% beta.hat.indep
  # imp.pls.indep = pls.loadings %*% beta.hat.indep
  imp.spatrf.cv = rslt$spatrf.cv$var.imp
  imp.spatrf.pl = rslt$spatrf.pl$var.imp
  imp.rf = rslt$rf.imp[,2]
  return(imp = cbind(imp.pls, imp.pls.indep, imp.spatrf.cv, 
                     imp.spatrf.pl, imp.rf))
}

imp.ufp.primary = getImp(grid.ufp.primary, dat$ufp_primary)
imp.ufp.uw = getImp(grid.ufp.uw, dat$ufp_uw)
imp.bc.primary = getImp(grid.bc.primary, dat$bc_primary)
imp.bc.uw = getImp(grid.bc.uw, dat$bc_uw)

plotImp = function(impMat, tt){
  my.alpha = .3
  colnames(impMat) = c("Method PLS", "Method PLS-Indep", 
                       "Method SpatRF-CV", "Method SpatRF-PL", "Method RF")
  
  split_cov_name <- function(dt, cov) {
    dt <- suppressWarnings(dt %>%
                             rename(cov_full_name = cov) %>%
                             mutate(
                               buffer = substr(cov_full_name, nchar(cov_full_name)-4, nchar(cov_full_name)),
                               buffer = as.numeric(ifelse(!is.na(as.integer(buffer)), buffer, NA)),
                               cov = ifelse(is.na(buffer), cov_full_name, substr(cov_full_name, 1, nchar(cov_full_name)-5) )
                             )
    ) %>%
      select(contains("cov"), buffer, everything())
    
    # elevation
    dt$cov[grepl("^elev_.+_above$", dt$cov_full_name)] <- "elev_above"
    dt$cov[grepl("^elev_.+_below$", dt$cov_full_name)] <- "elev_below"
    dt$cov[grepl("^elev_.+_stdev$", dt$cov_full_name)] <- "elev_stdev"
    dt$cov[grepl("^elev_.+_at_elev$", dt$cov_full_name)] <- "elev_at_elev"
    
    dt$buffer[grepl("_1k_", dt$cov_full_name)] <- 1000
    dt$buffer[grepl("_5k_", dt$cov_full_name)] <- 5000
    
    return(dt)
  }
  
  impMat = as.matrix(impMat)
  rownames(impMat) = colnames(X)
  dtNew <- impMat %>%
    as.data.frame() %>%
    rownames_to_column(var = "cov") %>%
    # rename variables if buffers
    split_cov_name(cov = "cov") %>%
      #make long format for faceting
      gather(key = "Method", value = "Importance", contains("Method")) %>%
      mutate(Method = substr(Method, 8, nchar(Method))) 
  #colnames(dtNew)[4] = "Importance"
  
  dtNew  %>%
    #buffered covariates
    drop_na(buffer) %>%
    ggplot(aes(x = Importance, y = cov)) +
    geom_point(aes(size=buffer),
               shape=1,
               alpha=my.alpha) +
    scale_size(breaks = c(min(dtNew$buffer, na.rm = T),
                          max(dtNew$buffer, na.rm = T)
    )) + #500, 5000, 10000,
    #non-buffered covariates
    geom_point(data = dtNew[is.na(dtNew$buffer),],
               alpha=my.alpha,
               aes(shape="")) +
    geom_vline(xintercept=0,
               linetype="solid",
               alpha=my.alpha) +
    facet_wrap(~ Method, scales = "free_x",
                 labeller = "label_both",
                 nrow = 1) +
    labs(y = "Geocovariate",
         shape= "non-buffer", #"proximity,\nelevation",
         #title = paste("PLS Importance:", pollutant)) +
         title = tt) +
    theme(legend.position = "bottom", text = element_text(size = 16))
}


#plotImp(imp.ufp.primary, "UFP (adjusted)")
plotImp(imp.ufp.uw, "UFP (unweighted)")
#plotImp(imp.bc.primary, "BC (adjusted)")
plotImp(imp.bc.uw, "BC (unweighted)")
```

### Pairwise scatterplot of variable importance

Absolute values were taken for PLS importance measures.

```{r}
# pairs(abs(imp.ufp.primary), pch = 19, main = "VIP: UFP (adjusted)",
#       labels = c("PLS", "PLS-indep", "SpatRF-CV", "SpatRF-PL", "RF"))
pairs(abs(imp.ufp.uw), pch = 19, main = "VIP: UFP (unweighted)",
      labels = c("PLS", "PLS-indep", "SpatRF-CV", "SpatRF-PL", "RF"))
# pairs(abs(imp.bc.primary), pch = 19, main = "VIP: BC (adjusted)",
#       labels = c("PLS", "PLS-indep", "SpatRF-CV", "SpatRF-PL", "RF"))
pairs(abs(imp.bc.uw), pch = 19, main = "VIP: BC (unweighted)",
      labels = c("PLS", "PLS-indep", "SpatRF-CV", "SpatRF-PL", "RF"))
```
