---
title: "(New) Variable Importance Summary, Natural Scale"
author: "Si Cheng"
date: ' `r Sys.Date()` '
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

## Approach 2: Changing predictor value from median

We consider an alternative approach proposed during the group meeting on Oct 11. The idea was to fix each predictor at their median (one at a time), and see how the predicted values change after increasing/decreasing them by a certain magnitude.

The model was trained with all 309 locations from the mobile monitoring data, and evaluated at each of these locations in a leave-one-out manner. The key idea was that we use the same mean model, but incorporate a new covariance part when looking at each location. We compare the predicted values when each predictor takes their 25%, 50% and 75% quantile.

```{r, fig.height=12, fig.width=15}
library(tidyverse)

dat = readRDS("/Users/sicheng/Library/CloudStorage/OneDrive-UW/exposurepred/trap/dat.rda")
X = dat[,-(1:21)]

cols = 1:ncol(X)
qs = c(.25, .5, .75)

plotImp = function(impMat, tt){
  my.alpha = .3
  colnames(impMat) = c("Method PLS", "Method SpatRF")# , "Method RF", "Method RF-TPRS")
  # impMat = impMat[,c(1,2)]
  
  split_cov_name <- function(dt, cov) {
    dt <- suppressWarnings(dt %>%
                             rename(cov_full_name = cov) %>%
                             mutate(
                               buffer = substr(cov_full_name, nchar(cov_full_name)-4, nchar(cov_full_name)),
                               buffer = as.numeric(ifelse(!is.na(as.integer(buffer)), buffer, NA)),
                               cov = ifelse(is.na(buffer), cov_full_name, substr(cov_full_name, 1, nchar(cov_full_name)-5) )
                             )
    ) %>%
      select(contains("cov"), buffer, everything())
    
    # elevation
    dt$cov[grepl("^elev_.+_above$", dt$cov_full_name)] <- "elev_above"
    dt$cov[grepl("^elev_.+_below$", dt$cov_full_name)] <- "elev_below"
    dt$cov[grepl("^elev_.+_stdev$", dt$cov_full_name)] <- "elev_stdev"
    dt$cov[grepl("^elev_.+_at_elev$", dt$cov_full_name)] <- "elev_at_elev"
    
    dt$buffer[grepl("_1k_", dt$cov_full_name)] <- 1000
    dt$buffer[grepl("_5k_", dt$cov_full_name)] <- 5000
    
    return(dt)
  }
  
  impMat = as.matrix(impMat)
  rownames(impMat) = colnames(X)
  dtNew <- impMat %>%
    as.data.frame() %>%
    rownames_to_column(var = "cov") %>%
    # rename variables if buffers
    split_cov_name(cov = "cov") %>%
      #make long format for faceting
      gather(key = "Method", value = "Importance", contains("Method")) %>%
      mutate(Method = substr(Method, 8, nchar(Method))) 
  #colnames(dtNew)[4] = "Importance"
  
  dtNew  %>%
    #buffered covariates
    drop_na(buffer) %>%
    ggplot(aes(x = Importance, y = cov)) +
    geom_point(aes(size=buffer),
               shape=1,
               alpha=my.alpha) +
    scale_size(breaks = c(min(dtNew$buffer, na.rm = T),
                          max(dtNew$buffer, na.rm = T)
    )) + #500, 5000, 10000,
    #non-buffered covariates
    geom_point(data = dtNew[is.na(dtNew$buffer),],
               alpha=my.alpha,
               aes(shape="")) +
    geom_vline(xintercept=0,
               linetype="solid",
               alpha=my.alpha) +
    facet_wrap(~ Method, scales = "fixed", # scales = "free_x",
                 labeller = "label_both",
                 nrow = 1) +
    labs(y = "Geocovariate", x = "Predictor Contribution",
         shape= "non-buffer", #"proximity,\nelevation",
         #title = paste("PLS Importance:", pollutant)) +
         title = tt) +
    theme(legend.position = "bottom", text = element_text(size = 16))
}
```

### UFP

```{r, echo = FALSE, fig.width=16, fig.height=15}

load("/Users/sicheng/Library/CloudStorage/OneDrive-UW/exposurepred/trap/varimp_new/varimp_loo_ufp_bc_updated.RData")

avg.pls = avg.spatrf = avg.rf = avg.rf.tprs = matrix(NA, ncol = length(qs), nrow = ncol(X))
k = 1

for (i in 1:length(cols)){
  for (j in 1:length(qs)){
    avg.pls[i,j] = mean(exp(rslt.ufp.uw$Y.pls[,k]))
    avg.spatrf[i,j] = mean(exp(rslt.ufp.uw$Y.spatrf.cv[,k]))
    k = k + 1
  }
}

plotImp(cbind((avg.pls[,2]) - (avg.pls[,1]), 
              (avg.spatrf[,2]) - (avg.spatrf[,1])),
        "Mean difference in prediction, 25% to 50% quantile, UFP")

plotImp(cbind((avg.pls[,3]) - (avg.pls[,2]), 
              (avg.spatrf[,3]) - (avg.spatrf[,2])),
        "Mean difference in prediction, 50% to 75% quantile, UFP")

plotImp(cbind((avg.pls[,3]) - (avg.pls[,1]), 
              (avg.spatrf[,3]) - (avg.spatrf[,1])),
        "Mean difference in prediction, 25% to 75% quantile, UFP")
```

The greatest differences between UK-PLS and SpatRF:

```{r}
vim_25_50 = cbind((avg.pls[,2]) - (avg.pls[,1]), 
              (avg.spatrf[,2]) - (avg.spatrf[,1])) %>% #scale() %>%
  as.data.frame() %>% mutate(diff = abs(V2 - V1), varname = colnames(X))
vim_50_75 = cbind((avg.pls[,3]) - (avg.pls[,2]), 
              (avg.spatrf[,3]) - (avg.spatrf[,2])) %>% #scale() %>%
  as.data.frame() %>% mutate(diff = abs(V2 - V1), varname = colnames(X))
vim_25_75 = cbind((avg.pls[,3]) - (avg.pls[,1]), 
              (avg.spatrf[,3]) - (avg.spatrf[,1])) %>% #scale() %>%
  as.data.frame() %>% mutate(diff = abs(V2 - V1), varname = colnames(X))

vim_25_50 %>% top_n(diff, n = 5) %>% 
  knitr::kable(col.names = c("PLS", "SpatRF", "abs(difference)", "Variable"),
               caption = "25% to 50% quantile, UFP")
vim_50_75 %>% top_n(diff, n = 5)%>% 
  knitr::kable(col.names = c("PLS", "SpatRF", "abs(difference)", "Variable"),
               caption = "50% to 75% quantile, UFP")
vim_25_75 %>% top_n(diff, n = 5) %>% 
  knitr::kable(col.names = c("PLS", "SpatRF", "abs(difference)", "Variable"),
               caption = "25% to 75% quantile, UFP")
```

Results on predictors corresponding to the greatest differences did not greatly differ from our previous version of variable importance evaluated with cohort data. Again, it seems that SpatRF is not capturing the impact of airport on UFP concentration, which was likely explained by other predictors instead.

It is also worth noting that (as in the last version) the sign of some predictor effects from SpatRF are different from UK-PLS (as well as our common sense), though their magnitude is often not large (see, e.g. distance to large airports, length of truck routes with buffer size 10000). Again, this is likely because such effects are captured by other predictors. 

**Cleaned variable importance plot with top 10 differences**

```{r, echo = FALSE, fig.width=16, fig.height=6}

plotImp2 = function(impMat1, impMat2, impMat3, tt, n_top = 5){
  # my.alpha = .3
  my.alpha = .8
  
  colnames(impMat1) = colnames(impMat2) = colnames(impMat3) =
    c("Method UK-PLS", "Method SpatRF")
  
  split_cov_name <- function(dt, cov) {
    dt <- suppressWarnings(dt %>%
                             rename(cov_full_name = cov) %>%
                             mutate(
                               buffer = substr(cov_full_name, nchar(cov_full_name)-4, nchar(cov_full_name)),
                               buffer = as.numeric(ifelse(!is.na(as.integer(buffer)), buffer, NA)),
                               cov = ifelse(is.na(buffer), cov_full_name, substr(cov_full_name, 1, nchar(cov_full_name)-5) )
                             )
    ) %>%
      select(contains("cov"), buffer, everything())
    
    # elevation
    dt$cov[grepl("^elev_.+_above$", dt$cov_full_name)] <- "elev_above"
    dt$cov[grepl("^elev_.+_below$", dt$cov_full_name)] <- "elev_below"
    dt$cov[grepl("^elev_.+_stdev$", dt$cov_full_name)] <- "elev_stdev"
    dt$cov[grepl("^elev_.+_at_elev$", dt$cov_full_name)] <- "elev_at_elev"
    
    dt$buffer[grepl("_1k_", dt$cov_full_name)] <- 1000
    dt$buffer[grepl("_5k_", dt$cov_full_name)] <- 5000
    
    return(dt)
  }
  
  impMat = rbind(impMat1, impMat2, impMat3) %>% as.data.frame()
  # impMat.std = rbind(impMat1.std, impMat2.std, impMat3.std) %>% as.data.frame()
  impMat$Quantiles = rep(c("25% to 50%", "50% to 75%", "25% to 75%"), 
                         each = nrow(impMat1))
  impMat$Diff = abs(impMat[,1] - impMat[,2])
  # impMat$Diff = abs(impMat.std[,1] - impMat.std[,2])
  impMat$cov = rep(colnames(X), 3)
  
  # impMat_top = impMat %>% group_by(Quantiles) %>% 
  #   slice_max(order_by = Diff, n = 5)
  
  #rownames(impMat) = rep(colnames(X), 3)
  dtNew <- impMat %>%
    as.data.frame() %>%
    #rownames_to_column(var = "cov") %>%
    # rename variables if buffers
    split_cov_name(cov = "cov") %>%
    #make long format for faceting
    gather(key = "Method", value = "Importance", contains("Method")) %>%
    mutate(Method = substr(Method, 8, nchar(Method))) 
  
  # set1 = dtNew %>% filter(Quantiles == "25% to 50%" & Method == "PLS") %>%
  #   slice_max(order_by = Diff, n = n_top) %>% select(cov) %>% c() %>% unlist()
  # set2 = dtNew %>% filter(Quantiles == "50% to 75%" & Method == "PLS") %>%
  #   slice_max(order_by = Diff, n = n_top) %>% select(cov) %>% c() %>% unlist()
  # set3 = dtNew %>% filter(Quantiles == "25% to 75%" & Method == "PLS") %>%
  #   slice_max(order_by = Diff, n = n_top) %>% select(cov) %>% c() %>% unlist()
  set1 = dtNew %>% filter(Quantiles == "25% to 50%" & Method == "UK-PLS") %>%
    slice_max(order_by = abs(Importance), n = n_top) %>% select(cov) %>% c() %>% unlist()
  set2 = dtNew %>% filter(Quantiles == "50% to 75%" & Method == "UK-PLS") %>%
    slice_max(order_by = abs(Importance), n = n_top) %>% select(cov) %>% c() %>% unlist()
  set3 = dtNew %>% filter(Quantiles == "25% to 75%" & Method == "UK-PLS") %>%
    slice_max(order_by = abs(Importance), n = n_top) %>% select(cov) %>% c() %>% unlist()
  set1.2 = dtNew %>% filter(Quantiles == "25% to 50%" & Method == "SpatRF") %>%
    slice_max(order_by = abs(Importance), n = n_top) %>% select(cov) %>% c() %>% unlist()
  set2.2 = dtNew %>% filter(Quantiles == "50% to 75%" & Method == "SpatRF") %>%
    slice_max(order_by = abs(Importance), n = n_top) %>% select(cov) %>% c() %>% unlist()
  set3.2 = dtNew %>% filter(Quantiles == "25% to 75%" & Method == "SpatRF") %>%
    slice_max(order_by = abs(Importance), n = n_top) %>% select(cov) %>% c() %>% unlist()
  # subset = set1 %>% union(set2) %>% union(set3)
  subset = set1 %>% union(set2) %>% union(set3) %>% 
    union(set1.2) %>% union(set2.2) %>% union(set3.2)
  dtNew = dtNew[dtNew$cov %in% subset, ] %>% arrange(cov)
  
  dtNew %>%
    #buffered covariates
    drop_na(buffer) %>%
    ggplot(aes(x = Importance, y = cov)) +
    geom_point(aes(size=buffer, color = Method), shape=1, alpha=my.alpha) +
    scale_size(breaks = c(min(dtNew$buffer, na.rm = T),
                          max(dtNew$buffer, na.rm = T))) + #500, 5000, 10000,
    #non-buffered covariates
    geom_point(data = dtNew[is.na(dtNew$buffer),],
               alpha=my.alpha, aes(shape="", color = Method)) +
    geom_vline(xintercept=0,
               linetype="solid",
               alpha=.3) +
    scale_y_discrete(labels=c("pop_s" = "Population", "ndvi_summer_a" = "NDVI (summer)",
                              "ndvi_q75_a" = "NDVI (annual 75%ile)",
                              "m_to_ry" = "Dist to railyard",
                              "m_to_l_airp" = "Dist to large airport",
                              "m_to_airp" = "Dist to airport",
                              "lu_transport_p" = "Transportation",
                              "lu_resi_p" = "Residential",
                              "lu_industrial_p" = "Industrial",
                              "lu_comm_p" = "Commercial",
                              "ll_a3_s" = "A3 road length",
                              "tl_s" = "Truck route length",
                              "rlu_evergreen_p" = "Evergreen forest",
                              "rlu_water_p" = "Water",
                              "pop10_s" = "Population",
                              "ll_a1_s" = "A1 road length",
                              "ll_a23_s" = "A2,A3 road length",
                              "rlu_dev_hi_p" = "Dev high intensity",
                              "rlu_dev_med_p" = "Dev medium intensity",
                              "log_m_to_ry" = "Log(dist) to railyard",
                              "log_m_to_l_airp" = "Log(dist) to large airport",
                              "log_m_to_airp" = "Log(dist) to airport",
                              "log_m_to_m_port" = "Log(dist) to medium ports",
                              "log_m_to_a1" = "Log(dist) to A1 road",
                              "log_m_to_a123" = "Log(dist) to A1,A2,A3 road",
                              "log_m_to_a1_a1_intersect" = "Log(dist) to A1/A1 intersec",
                              "log_m_to_a2_a3_intersect" = "Log(dist) to A2/A3 intersec",
                              "log_m_to_a2_a2_intersect" = "Log(dist) to A2/A2 intersec",
                              "log_m_to_a1_a3_intersect" = "Log(dist) to A1/A3 intersec",
                              "log_m_to_truck" = "Log(dist) to truck route",
                              "log_m_to_waterway" = "Log(dist) to waterway",
                              "log_m_to_rr" = "Log(dist) to railroad",
                              "imp_a" = "Impervious surface",
                              "elev_elevation" = "Elevation",
                              "elev_below" = "High relative elevation",
                              "elev_at_elev" = "Intermediate relative elevation",
                              "elev_above" = "Low relative elevation")) +
    facet_wrap(~ Quantiles, scales = "fixed", # scales = "free_x",
               labeller = "label_both",
               nrow = 1) +
    labs(y = "Geocovariate", x = "Predictor Contribution",
         shape= "non-buffer", #"proximity,\nelevation",
         #title = paste("PLS Importance:", pollutant)) +
         title = tt) +
    theme(legend.position = "bottom", text = element_text(size = 16))
}

plotImp2(cbind((avg.pls[,2]) - (avg.pls[,1]), 
              (avg.spatrf[,2]) - (avg.spatrf[,1])),
         cbind((avg.pls[,3]) - (avg.pls[,2]), 
              (avg.spatrf[,3]) - (avg.spatrf[,2])),
         cbind((avg.pls[,3]) - (avg.pls[,1]), 
              (avg.spatrf[,3]) - (avg.spatrf[,1])),
        "Predictors with Top Contribution, UFP", n_top = 5)
```

### BC

```{r, echo = FALSE, fig.width=16, fig.height=15}

avg.pls = avg.spatrf = avg.rf = avg.rf.tprs = matrix(NA, ncol = length(qs), nrow = ncol(X))
k = 1

for (i in 1:length(cols)){
  for (j in 1:length(qs)){
    avg.pls[i,j] = mean(exp(rslt.bc.uw$Y.pls[,k]))
    avg.spatrf[i,j] = mean(exp(rslt.bc.uw$Y.spatrf.cv[,k]))
    k = k + 1
  }
}

plotImp(cbind((avg.pls[,2]) - (avg.pls[,1]), 
              (avg.spatrf[,2]) - (avg.spatrf[,1])),
        "Mean difference in prediction, 25% to 50% quantile, BC")

plotImp(cbind((avg.pls[,3]) - (avg.pls[,2]), 
              (avg.spatrf[,3]) - (avg.spatrf[,2])),
        "Mean difference in prediction, 50% to 75% quantile, BC")

plotImp(cbind((avg.pls[,3]) - (avg.pls[,1]), 
              (avg.spatrf[,3]) - (avg.spatrf[,1])),
        "Mean difference in prediction, 25% to 75% quantile, BC")
```

The greatest differences between UK-PLS and SpatRF:

```{r}
vim_25_50 = cbind((avg.pls[,2]) - (avg.pls[,1]), 
              (avg.spatrf[,2]) - (avg.spatrf[,1])) %>% #scale() %>%
  as.data.frame() %>% mutate(diff = abs(V2 - V1), varname = colnames(X))
vim_50_75 = cbind((avg.pls[,3]) - (avg.pls[,2]), 
              (avg.spatrf[,3]) - (avg.spatrf[,2])) %>% #scale() %>%
  as.data.frame() %>% mutate(diff = abs(V2 - V1), varname = colnames(X))
vim_25_75 = cbind((avg.pls[,3]) - (avg.pls[,1]), 
              (avg.spatrf[,3]) - (avg.spatrf[,1])) %>% #scale() %>%
  as.data.frame() %>% mutate(diff = abs(V2 - V1), varname = colnames(X))

vim_25_50 %>% top_n(diff, n = 5) %>% 
  knitr::kable(col.names = c("PLS", "SpatRF", "abs(difference)", "Variable"),
               caption = "25% to 50% quantile, BC")
vim_50_75 %>% top_n(diff, n = 5)%>% 
  knitr::kable(col.names = c("PLS", "SpatRF", "abs(difference)", "Variable"),
               caption = "50% to 75% quantile, BC")
vim_25_75 %>% top_n(diff, n = 5) %>% 
  knitr::kable(col.names = c("PLS", "SpatRF", "abs(difference)", "Variable"),
               caption = "25% to 75% quantile, BC")
```

Summarizing in words, the predictors for which the importance differs the most are

* 25% to 50%: Population, distance to large airports, distance to rail yards, length of truck routes, imperviousness
* 50% to 75%: Length of truck routes, evergreen forest land, population
* 25% to 75%: Length of truck routes, evergreen forest land, population

Comparing to previous results on cohort data, (relative) elevation did not appear among the top 5 predictors on which PLS and SpatRF variable importance differs. However, it was the 6th or 7th in these new results, which does not seem to be drastically different than before.

**Cleaned variable importance plot with top 10 differences**

```{r, fig.width = 16, fig.height = 6}
plotImp2(cbind((avg.pls[,2]) - (avg.pls[,1]), 
              (avg.spatrf[,2]) - (avg.spatrf[,1])),
         cbind((avg.pls[,3]) - (avg.pls[,2]), 
              (avg.spatrf[,3]) - (avg.spatrf[,2])),
         cbind((avg.pls[,3]) - (avg.pls[,1]), 
              (avg.spatrf[,3]) - (avg.spatrf[,1])),
        "Predictors with Top Contribution, BC", n_top = 5)
```

### NO2

```{r, echo = FALSE, fig.width=16, fig.height=15}
load("Users/sicheng/Library/CloudStorage/OneDrive-UW/exposurepred/trap/varimp_new/varimp_loo_xo2.RData")

avg.pls = avg.spatrf = avg.rf = avg.rf.tprs = matrix(NA, ncol = length(qs), nrow = ncol(X))
k = 1

for (i in 1:length(cols)){
  for (j in 1:length(qs)){
    avg.pls[i,j] = mean(exp(rslt.no2.uw$Y.pls[,k]))
    avg.spatrf[i,j] = mean(exp(rslt.no2.uw$Y.spatrf.cv[,k]))
    k = k + 1
  }
}

plotImp(cbind((avg.pls[,2]) - (avg.pls[,1]), 
              (avg.spatrf[,2]) - (avg.spatrf[,1])),
        "Mean difference in prediction, 25% to 50% quantile, NO2")

plotImp(cbind((avg.pls[,3]) - (avg.pls[,2]), 
              (avg.spatrf[,3]) - (avg.spatrf[,2])),
        "Mean difference in prediction, 50% to 75% quantile, NO2")

plotImp(cbind((avg.pls[,3]) - (avg.pls[,1]), 
              (avg.spatrf[,3]) - (avg.spatrf[,1])),
        "Mean difference in prediction, 25% to 75% quantile, NO2")
```

The greatest differences between UK-PLS and SpatRF:

```{r}
vim_25_50 = cbind((avg.pls[,2]) - (avg.pls[,1]), 
              (avg.spatrf[,2]) - (avg.spatrf[,1])) %>% #scale() %>%
  as.data.frame() %>% mutate(diff = abs(V2 - V1), varname = colnames(X))
vim_50_75 = cbind((avg.pls[,3]) - (avg.pls[,2]), 
              (avg.spatrf[,3]) - (avg.spatrf[,2])) %>% #scale() %>%
  as.data.frame() %>% mutate(diff = abs(V2 - V1), varname = colnames(X))
vim_25_75 = cbind((avg.pls[,3]) - (avg.pls[,1]), 
              (avg.spatrf[,3]) - (avg.spatrf[,1])) %>% #scale() %>%
  as.data.frame() %>% mutate(diff = abs(V2 - V1), varname = colnames(X))

vim_25_50 %>% top_n(diff, n = 5) %>% 
  knitr::kable(col.names = c("PLS", "SpatRF", "abs(difference)", "Variable"),
               caption = "25% to 50% quantile, NO2")
vim_50_75 %>% top_n(diff, n = 5)%>% 
  knitr::kable(col.names = c("PLS", "SpatRF", "abs(difference)", "Variable"),
               caption = "50% to 75% quantile, NO2")
vim_25_75 %>% top_n(diff, n = 5) %>% 
  knitr::kable(col.names = c("PLS", "SpatRF", "abs(difference)", "Variable"),
               caption = "25% to 75% quantile, NO2")
```

**Cleaned variable importance plot with top 10 differences**

```{r, fig.width = 16, fig.height = 6}
plotImp2(cbind((avg.pls[,2]) - (avg.pls[,1]), 
              (avg.spatrf[,2]) - (avg.spatrf[,1])),
         cbind((avg.pls[,3]) - (avg.pls[,2]), 
              (avg.spatrf[,3]) - (avg.spatrf[,2])),
         cbind((avg.pls[,3]) - (avg.pls[,1]), 
              (avg.spatrf[,3]) - (avg.spatrf[,1])),
        "Predictors with Top Contribution, NO2", n_top = 5)
```

### CO2

```{r, echo = FALSE, fig.width=16, fig.height=15}

avg.pls = avg.spatrf = avg.rf = avg.rf.tprs = matrix(NA, ncol = length(qs), nrow = ncol(X))
k = 1

for (i in 1:length(cols)){
  for (j in 1:length(qs)){
    avg.pls[i,j] = mean(exp(rslt.co2.uw$Y.pls[,k]))
    avg.spatrf[i,j] = mean(exp(rslt.co2.uw$Y.spatrf.cv[,k]))
    k = k + 1
  }
}

plotImp(cbind((avg.pls[,2]) - (avg.pls[,1]), 
              (avg.spatrf[,2]) - (avg.spatrf[,1])),
        "Mean difference in prediction, 25% to 50% quantile, CO2")

plotImp(cbind((avg.pls[,3]) - (avg.pls[,2]), 
              (avg.spatrf[,3]) - (avg.spatrf[,2])),
        "Mean difference in prediction, 50% to 75% quantile, CO2")

plotImp(cbind((avg.pls[,3]) - (avg.pls[,1]), 
              (avg.spatrf[,3]) - (avg.spatrf[,1])),
        "Mean difference in prediction, 25% to 75% quantile, CO2")
```

The greatest differences between UK-PLS and SpatRF:

```{r}
vim_25_50 = cbind((avg.pls[,2]) - (avg.pls[,1]), 
              (avg.spatrf[,2]) - (avg.spatrf[,1])) %>% #scale() %>%
  as.data.frame() %>% mutate(diff = abs(V2 - V1), varname = colnames(X))
vim_50_75 = cbind((avg.pls[,3]) - (avg.pls[,2]), 
              (avg.spatrf[,3]) - (avg.spatrf[,2])) %>% #scale() %>%
  as.data.frame() %>% mutate(diff = abs(V2 - V1), varname = colnames(X))
vim_25_75 = cbind((avg.pls[,3]) - (avg.pls[,1]), 
              (avg.spatrf[,3]) - (avg.spatrf[,1])) %>% #scale() %>%
  as.data.frame() %>% mutate(diff = abs(V2 - V1), varname = colnames(X))

vim_25_50 %>% top_n(diff, n = 5) %>% 
  knitr::kable(col.names = c("PLS", "SpatRF", "abs(difference)", "Variable"),
               caption = "25% to 50% quantile, CO2")
vim_50_75 %>% top_n(diff, n = 5)%>% 
  knitr::kable(col.names = c("PLS", "SpatRF", "abs(difference)", "Variable"),
               caption = "50% to 75% quantile, CO2")
vim_25_75 %>% top_n(diff, n = 5) %>% 
  knitr::kable(col.names = c("PLS", "SpatRF", "abs(difference)", "Variable"),
               caption = "25% to 75% quantile, CO2")
```

**Cleaned variable importance plot with top 10 differences**

```{r, fig.width = 16, fig.height = 6}
plotImp2(cbind((avg.pls[,2]) - (avg.pls[,1]), 
              (avg.spatrf[,2]) - (avg.spatrf[,1])),
         cbind((avg.pls[,3]) - (avg.pls[,2]), 
              (avg.spatrf[,3]) - (avg.spatrf[,2])),
         cbind((avg.pls[,3]) - (avg.pls[,1]), 
              (avg.spatrf[,3]) - (avg.spatrf[,1])),
        "Predictors with Top Contribution, CO2", n_top = 5)
```

### PM2.5

```{r, echo = FALSE, fig.width=16, fig.height=15}

load("C:/Users/Si Cheng/Dropbox (Personal)/exposurepred/trap/varimp_new/varimp_loo_pm25.RData")

avg.pls = avg.spatrf = avg.rf = avg.rf.tprs = matrix(NA, ncol = length(qs), nrow = ncol(X))
k = 1

for (i in 1:length(cols)){
  for (j in 1:length(qs)){
    avg.pls[i,j] = mean(exp(rslt.pm25.uw$Y.pls[,k]))
    avg.spatrf[i,j] = mean(exp(rslt.pm25.uw$Y.spatrf.cv[,k]))
    k = k + 1
  }
}

plotImp(cbind((avg.pls[,2]) - (avg.pls[,1]), 
              (avg.spatrf[,2]) - (avg.spatrf[,1])),
        "Mean difference in prediction, 25% to 50% quantile, PM2.5")

plotImp(cbind((avg.pls[,3]) - (avg.pls[,2]), 
              (avg.spatrf[,3]) - (avg.spatrf[,2])),
        "Mean difference in prediction, 50% to 75% quantile, PM2.5")

plotImp(cbind((avg.pls[,3]) - (avg.pls[,1]), 
              (avg.spatrf[,3]) - (avg.spatrf[,1])),
        "Mean difference in prediction, 25% to 75% quantile, PM2.5")
```

The greatest differences between UK-PLS and SpatRF:

```{r}
vim_25_50 = cbind((avg.pls[,2]) - (avg.pls[,1]), 
              (avg.spatrf[,2]) - (avg.spatrf[,1])) %>% #scale() %>%
  as.data.frame() %>% mutate(diff = abs(V2 - V1), varname = colnames(X))
vim_50_75 = cbind((avg.pls[,3]) - (avg.pls[,2]), 
              (avg.spatrf[,3]) - (avg.spatrf[,2])) %>% #scale() %>%
  as.data.frame() %>% mutate(diff = abs(V2 - V1), varname = colnames(X))
vim_25_75 = cbind((avg.pls[,3]) - (avg.pls[,1]), 
              (avg.spatrf[,3]) - (avg.spatrf[,1])) %>% #scale() %>%
  as.data.frame() %>% mutate(diff = abs(V2 - V1), varname = colnames(X))

vim_25_50 %>% top_n(diff, n = 5) %>% 
  knitr::kable(col.names = c("PLS", "SpatRF", "abs(difference)", "Variable"),
               caption = "25% to 50% quantile, PM2.5")
vim_50_75 %>% top_n(diff, n = 5)%>% 
  knitr::kable(col.names = c("PLS", "SpatRF", "abs(difference)", "Variable"),
               caption = "50% to 75% quantile, PM2.5")
vim_25_75 %>% top_n(diff, n = 5) %>% 
  knitr::kable(col.names = c("PLS", "SpatRF", "abs(difference)", "Variable"),
               caption = "25% to 75% quantile, PM2.5")
```

**Cleaned variable importance plot with top 10 differences**

```{r, fig.width = 16, fig.height = 6}
plotImp2(cbind((avg.pls[,2]) - (avg.pls[,1]), 
              (avg.spatrf[,2]) - (avg.spatrf[,1])),
         cbind((avg.pls[,3]) - (avg.pls[,2]), 
              (avg.spatrf[,3]) - (avg.spatrf[,2])),
         cbind((avg.pls[,3]) - (avg.pls[,1]), 
              (avg.spatrf[,3]) - (avg.spatrf[,1])),
        "Predictors with Top Contribution, PM2.5", n_top = 5)
```