---
title: "Modeling PM2.5 sub-species concentration"
author: "Si Cheng"
date: ' `r Sys.Date()` '
output: 
  html_document:
    number_sections: no
    df_print: paged
    toc: yes
    toc_depth: '3'

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message = FALSE)

load("/Users/sicheng/Library/CloudStorage/OneDrive-UW/exposurepred/spatialRF/3_ec_all_full.RData")
rslt.ec = rslt
annavg.ec = annavg
plltnt.ec = plltnt.mat

load("/Users/sicheng/Library/CloudStorage/OneDrive-UW/exposurepred/spatialRF/3_oc_all.RData")
rslt.oc = rslt
annavg.oc = annavg

load("/Users/sicheng/Library/CloudStorage/OneDrive-UW/exposurepred/spatialRF/3_s_all_full.RData")
rslt.s = rslt
annavg.s = annavg
plltnt.s = plltnt.mat
load("/Users/sicheng/Library/CloudStorage/OneDrive-UW/exposurepred/spatialRF/3_si_all.RData")
rslt.si = rslt
annavg.si = annavg
rm(rslt, annavg)
```

## Overview
This summary includes national air pollution models for PM2.5 sub-species, with a focus on variable importance measures. The tables below show cross-validated $R^2$ for UK-PLS and spatial RF (on the native and square root scale, respectively).

```{r, echo = FALSE}
# combine predictions
time.pls <- time.spatrf <- 0
Y.pls.oc = Y.spatrf.cv.oc = Y.spatrf.pl.oc = Y.rf.oc = Y.tprs.oc = Y.rf.tprs.oc = Y.tprs.rf.oc =
  Y.pls = Y.spatrf.cv = Y.spatrf.pl = Y.rf = Y.tprs = Y.rf.tprs = Y.tprs.rf = 
  rep(NA, length(rslt.ec[[1]]$Y.pls))
Y.pls.s = Y.spatrf.cv.s = Y.spatrf.pl.s = Y.rf.s = Y.tprs.s = Y.rf.tprs.s = Y.tprs.rf.s =
  Y.pls.si = Y.spatrf.cv.si = Y.spatrf.pl.si = Y.rf.si = Y.tprs.si = Y.rf.tprs.si = Y.tprs.rf.si =
  rep(NA, length(rslt.s[[1]]$Y.pls))

for (j in 1:length(rslt.ec)){
  idx <- !is.na(rslt.ec[[j]]$Y.pls)
  Y.pls[idx] <- rslt.ec[[j]]$Y.pls[idx]
  Y.spatrf.cv[idx] <- rslt.ec[[j]]$Y.spatrf.cv[idx]
  Y.spatrf.pl[idx] <- rslt.ec[[j]]$Y.spatrf.pl[idx]
  Y.rf[idx] = rslt.ec[[j]]$Y.rf[idx]
  Y.tprs[idx] = rslt.ec[[j]]$Y.tprs[idx]
  Y.rf.tprs[idx] = rslt.ec[[j]]$Y.rf.tprs[idx]
  Y.tprs.rf[idx] = rslt.ec[[j]]$Y.tprs.rf[idx]
  time.pls <- time.pls + rslt.ec[[j]]$pls.time[3]
  time.spatrf <- time.spatrf + rslt.ec[[j]]$spatrf.time[3]
  
  idx2 <- !is.na(rslt.oc[[j]]$Y.pls)
  Y.pls.oc[idx2] <- rslt.oc[[j]]$Y.pls[idx2]
  Y.spatrf.cv.oc[idx2] <- rslt.oc[[j]]$Y.spatrf.cv[idx2]
  Y.spatrf.pl.oc[idx2] <- rslt.oc[[j]]$Y.spatrf.pl[idx2]
  Y.rf.oc[idx2] = rslt.oc[[j]]$Y.rf[idx2]
  Y.tprs.oc[idx2] = rslt.oc[[j]]$Y.tprs[idx2]
  Y.rf.tprs.oc[idx2] = rslt.oc[[j]]$Y.rf.tprs[idx2]
  Y.tprs.rf.oc[idx2] = rslt.oc[[j]]$Y.tprs.rf[idx2]
  
  idx3 <- !is.na(rslt.s[[j]]$Y.pls)
  Y.pls.s[idx3] <- rslt.s[[j]]$Y.pls[idx3]
  Y.spatrf.cv.s[idx3] <- rslt.s[[j]]$Y.spatrf.cv[idx3]
  Y.spatrf.pl.s[idx3] <- rslt.s[[j]]$Y.spatrf.pl[idx3]
  Y.rf.s[idx3] = rslt.s[[j]]$Y.rf[idx3]
  Y.tprs.s[idx3] = rslt.s[[j]]$Y.tprs[idx3]
  Y.rf.tprs.s[idx3] = rslt.s[[j]]$Y.rf.tprs[idx3]
  Y.tprs.rf.s[idx3] = rslt.s[[j]]$Y.tprs.rf[idx3]
  
  idx4 <- !is.na(rslt.si[[j]]$Y.pls)
  Y.pls.si[idx4] <- rslt.si[[j]]$Y.pls[idx4]
  Y.spatrf.cv.si[idx4] <- rslt.si[[j]]$Y.spatrf.cv[idx4]
  Y.spatrf.pl.si[idx4] <- rslt.si[[j]]$Y.spatrf.pl[idx4]
  Y.rf.si[idx4] = rslt.si[[j]]$Y.rf[idx4]
  Y.tprs.si[idx4] = rslt.si[[j]]$Y.tprs[idx4]
  Y.rf.tprs.si[idx4] = rslt.si[[j]]$Y.rf.tprs[idx4]
  Y.tprs.rf.si[idx4] = rslt.si[[j]]$Y.tprs.rf[idx4]
}
Y.pls[Y.pls < 0] = 0
Y.pls.oc[Y.pls.oc < 0] = 0
Y.spatrf.cv.oc[Y.spatrf.cv.oc < 0] = 0
Y.spatrf.pl.oc[Y.spatrf.pl.oc < 0] = 0

Y.pls.s[Y.pls.s < 0] = 0
Y.spatrf.cv.s[Y.spatrf.cv.s < 0] = 0
Y.spatrf.pl.s[Y.spatrf.pl.s < 0] = 0

Y.pls.si[Y.pls.si < 0] = 0

r2.table = matrix(c(1 - mean((Y.pls^2 - annavg.ec^2)^2)/var(annavg.ec^2),
                    1 - mean((Y.spatrf.pl^2 - annavg.ec^2)^2)/var(annavg.ec^2),
                    1 - mean((Y.spatrf.cv^2 - annavg.ec^2)^2)/var(annavg.ec^2),
                    1 - mean((Y.rf^2 - annavg.ec^2)^2)/var(annavg.ec^2),
                    1 - mean((Y.tprs^2 - annavg.ec^2)^2)/var(annavg.ec^2),
                    1 - mean((Y.rf.tprs^2 - annavg.ec^2)^2)/var(annavg.ec^2),
                    1 - mean((Y.tprs.rf^2 - annavg.ec^2)^2)/var(annavg.ec^2),
                    .73, .83, .83, rep(NA, 4),
                    1 - mean((Y.pls.oc^2 - annavg.oc^2)^2)/var(annavg.oc^2),
                    1 - mean((Y.spatrf.pl.oc^2 - annavg.oc^2)^2)/var(annavg.oc^2),
                    1 - mean((Y.spatrf.cv.oc^2 - annavg.oc^2)^2)/var(annavg.oc^2),
                    1 - mean((Y.rf.oc^2 - annavg.oc^2)^2)/var(annavg.oc^2),
                    1 - mean((Y.tprs.oc^2 - annavg.oc^2)^2)/var(annavg.oc^2),
                    1 - mean((Y.rf.tprs.oc^2 - annavg.oc^2)^2)/var(annavg.oc^2),
                    1 - mean((Y.tprs.rf.oc^2 - annavg.oc^2)^2)/var(annavg.oc^2),
                    .58, .59, .63, rep(NA, 4),
                    1 - mean((Y.pls.s^2 - annavg.s^2)^2)/var(annavg.s^2),
                    1 - mean((Y.spatrf.pl.s^2 - annavg.s^2)^2)/var(annavg.s^2),
                    1 - mean((Y.spatrf.cv.s^2 - annavg.s^2)^2)/var(annavg.s^2),
                    1 - mean((Y.rf.s^2 - annavg.s^2)^2)/var(annavg.s^2),
                    1 - mean((Y.tprs.s^2 - annavg.s^2)^2)/var(annavg.s^2),
                    1 - mean((Y.rf.tprs.s^2 - annavg.s^2)^2)/var(annavg.s^2),
                    1 - mean((Y.tprs.rf.s^2 - annavg.s^2)^2)/var(annavg.s^2),
                    # 1 - mean((Y.pls.s[-257]^2 - annavg.s[-257]^2)^2)/var(annavg.s[-257]^2),
                    # 1 - mean((Y.spatrf.pl.s[-257]^2 - annavg.s[-257]^2)^2)/var(annavg.s[-257]^2),
                    # 1 - mean((Y.spatrf.cv.s[-257]^2 - annavg.s[-257]^2)^2)/var(annavg.s[-257]^2),
                    .95, .94, .94, rep(NA, 4),
                    1 - mean((Y.pls.si^2 - annavg.si^2)^2)/var(annavg.si^2),
                    1 - mean((Y.spatrf.pl.si^2 - annavg.si^2)^2)/var(annavg.si^2),
                    1 - mean((Y.spatrf.cv.si^2 - annavg.si^2)^2)/var(annavg.si^2),
                    1 - mean((Y.rf.si^2 - annavg.si^2)^2)/var(annavg.si^2),
                    1 - mean((Y.tprs.si^2 - annavg.si^2)^2)/var(annavg.si^2),
                    1 - mean((Y.rf.tprs.si^2 - annavg.si^2)^2)/var(annavg.si^2),
                    1 - mean((Y.tprs.rf.si^2 - annavg.si^2)^2)/var(annavg.si^2),
                    .56, .58, .60, rep(NA, 4)),
                  nrow = 8, byrow = TRUE)
colnames(r2.table) = c("UK-PLS", "SpatRF-PL", "SpatRF-CV", "RF", "TPRS", "RF-TPRS", "TPRS-RF")
rownames(r2.table) = c("EC", "(Original)", "OC", "(Original)", "S", # "S (one outlier removed)", 
                       "(Original)", "Si", "(Original)")

knitr::kable(r2.table, digits = 4, caption = "R2 on the native scale")

r2.table2 = matrix(c(1 - mean((Y.pls - annavg.ec)^2)/var(annavg.ec),
                    1 - mean((Y.spatrf.pl - annavg.ec)^2)/var(annavg.ec),
                    1 - mean((Y.spatrf.cv - annavg.ec)^2)/var(annavg.ec),
                    .82, NA, NA,
                    1 - mean((Y.pls.oc - annavg.oc)^2)/var(annavg.oc),
                    1 - mean((Y.spatrf.pl.oc - annavg.oc)^2)/var(annavg.oc),
                    1 - mean((Y.spatrf.cv.oc - annavg.oc)^2)/var(annavg.oc),
                    .69, NA, NA,
                    1 - mean((Y.pls.s - annavg.s)^2)/var(annavg.s),
                    1 - mean((Y.spatrf.pl.s - annavg.s)^2)/var(annavg.s),
                    1 - mean((Y.spatrf.cv.s - annavg.s)^2)/var(annavg.s),
                    .95, NA, NA,
                    1 - mean((Y.pls.si - annavg.si)^2)/var(annavg.si),
                    1 - mean((Y.spatrf.pl.si - annavg.si)^2)/var(annavg.si),
                    1 - mean((Y.spatrf.cv.si - annavg.si)^2)/var(annavg.si),
                    .62, NA, NA),
                  nrow = 8, byrow = TRUE)
colnames(r2.table2) = c("UK-PLS", "SpatRF-PL", "SpatRF-CV")
rownames(r2.table2) = c("EC", "(Bergen)", "OC", "(Bergen)",
                        "S", "(Bergen)", "Si", "(Bergen)")

knitr::kable(r2.table2, digits = 4, caption = "R2 on the square root scale")
```

## Prediction and residual plots
### EC

The figure below shows a comparison of predicted EC concentration from different models.

```{r, echo = FALSE, fig.width=24, fig.height=15}
library(ggmap)
library(ggplot2)
library(gridExtra)
library(ggpubr)
library(dplyr)

plot_pred = function(annavg, Y.pls, Y.spatrf.pl, Y.spatrf.cv, tt, plltnt.mat){
  rslt <- as.data.frame(cbind(annavg^2, Y.pls^2, Y.spatrf.pl^2, Y.spatrf.cv^2, 
                               plltnt.mat$longitude, plltnt.mat$latitude))
  colnames(rslt) <- c("Obs","UK_PLS", "SpatRF_PL", "SpatRF_CV", 
                       "longitude", "latitude")

  bbox = c(left = min(rslt$longitude) - 2, bottom = min(rslt$latitude) - 2,
         right = max(rslt$longitude) + 2, top = max(rslt$latitude) + 2)
  usMap = suppressMessages(get_stamenmap(
    bbox=bbox,
    zoom = 6,
    maptype = "toner-lite" #has airport symbol
  #maptype = "toner-background" #roads & water but no airport
  ))
  # usMap <- get_map(location=c(min(rslt$longitude), min(rslt$latitude),
  #                           max(rslt$longitude), max(rslt$latitude)),
  #                source="stamen", maptype = "terrain", crop=FALSE)

  p.obs <- ggmap(usMap, darken = c(.5, "white")) + 
         geom_point(data = rslt, 
                    aes(x = longitude, y = latitude, color = Obs, size = 1)) +
         scale_color_viridis_c(option = "plasma", name = "", trans = "sqrt",
                                  limits = range(rslt[,1:4])) +
         scale_x_continuous(limits = c(min(rslt$longitude)-2,
                                       max(rslt$longitude)+2), 
                            expand = c(0, 0)) +
         scale_y_continuous(limits = c(min(rslt$latitude)-2,
                                       max(rslt$latitude)+2), 
                            expand = c(0, 0)) +
         guides(size = FALSE) + ggtitle("Observed") +
    labs(x = "Longitude", y = "Latitude")
         
  p.pls <- ggmap(usMap, darken = c(.5, "white")) + 
         geom_point(data = rslt, 
                    aes(x = longitude, y = latitude, color = UK_PLS, size = 1)) +
         scale_color_viridis_c(option = "plasma", name = "", trans = "sqrt",
                                  limits = range(rslt[,1:4])) +
        scale_x_continuous(limits = c(min(rslt$longitude)-2,
                                       max(rslt$longitude)+2), 
                            expand = c(0, 0)) +
         scale_y_continuous(limits = c(min(rslt$latitude)-2,
                                       max(rslt$latitude)+2), 
                            expand = c(0, 0)) +
         guides(size = FALSE) + ggtitle("UK PLS") +
    labs(x = "Longitude", y = "Latitude")

  p.pl <- ggmap(usMap, darken = c(.5, "white")) + 
         geom_point(data = rslt, 
                    aes(x = longitude, y = latitude, 
                        color = SpatRF_PL, size = 1)) +
         scale_color_viridis_c(option = "plasma", name = "", trans = "sqrt",
                                  limits = range(rslt[,1:4])) +
        scale_x_continuous(limits = c(min(rslt$longitude)-2,
                                       max(rslt$longitude)+2), 
                            expand = c(0, 0)) +
         scale_y_continuous(limits = c(min(rslt$latitude)-2,
                                       max(rslt$latitude)+2), 
                            expand = c(0, 0)) +
        guides(size = FALSE) + ggtitle("SpatRF-PL") +
    labs(x = "Longitude", y = "Latitude")

  p.cv <- ggmap(usMap, darken = c(.5, "white")) + 
         geom_point(data = rslt, 
                    aes(x = longitude, y = latitude, 
                        color = SpatRF_CV, size = 1)) +
         scale_color_viridis_c(option = "plasma", name = "", trans = "sqrt",
                                  limits = range(rslt[,1:4])) +
        scale_x_continuous(limits = c(min(rslt$longitude)-2,
                                       max(rslt$longitude)+2), 
                            expand = c(0, 0)) +
         scale_y_continuous(limits = c(min(rslt$latitude)-2,
                                       max(rslt$latitude)+2), 
                            expand = c(0, 0)) +
        guides(size = FALSE) + ggtitle("SpatRF-CV") +
    labs(x = "Longitude", y = "Latitude")

  ggarrange(p.obs,p.pls,p.pl,p.cv,
            nrow = 2, ncol = 2, common.legend = TRUE, legend = "bottom") %>%
    annotate_figure(top = text_grob(tt, face = "bold", size = 16))
  # grid.arrange(arrangeGrob(p.obs,p.pls,p.pl,p.cv,nrow=2))
}

plot_pred(annavg.ec, Y.pls, Y.spatrf.pl, Y.spatrf.cv, "EC predictions", plltnt.ec)
```

Residual (predicted minus observed) map:

```{r, echo = FALSE, fig.width=24, fig.height=12}
# library(tidyverse)

plot_res = function(annavg, Y.pls, Y.spatrf.pl, Y.spatrf.cv, 
                    Y.rf, Y.tprs, Y.rf.tprs, Y.tprs.rf,
                    tt, plltnt.mat){
  rslt <- as.data.frame(cbind(annavg^2, Y.pls^2, Y.spatrf.pl^2, Y.spatrf.cv^2, 
                              Y.rf^2, Y.tprs^2, Y.rf.tprs^2, Y.tprs.rf^2,
                               plltnt.mat$longitude, plltnt.mat$latitude))
  colnames(rslt) <- c("Obs","UK_PLS", "SpatRF_PL", "SpatRF_CV", 
                      "RF", "TPRS", "RF_TPRS", "TPRS_RF",
                       "longitude", "latitude")
  
  rslt$res_pls = rslt$UK_PLS - rslt$Obs
  rslt$res_spatrf = rslt$SpatRF_PL - rslt$Obs
  rslt$res_rf = rslt$RF - rslt$Obs
  rslt$res_tprs = rslt$TPRS - rslt$Obs
  rslt$res_rf_tprs = rslt$RF_TPRS - rslt$Obs
  rslt$res_tprs_rf = rslt$TPRS_RF - rslt$Obs

  bbox = c(left = min(rslt$longitude) - 2, bottom = min(rslt$latitude) - 2,
         right = max(rslt$longitude) + 2, top = max(rslt$latitude) + 2)
  usMap = suppressMessages(get_stamenmap(
    bbox=bbox,
    zoom = 6,
    maptype = "toner-lite" #has airport symbol
  #maptype = "toner-background" #roads & water but no airport
  ))
  
  # usMap <- get_map(location=c(min(rslt$longitude), min(rslt$latitude),
  #                           max(rslt$longitude), max(rslt$latitude)),
  #                source="stamen", maptype = "terrain", crop=FALSE)

  p1.pls <- ggmap(usMap, darken = c(.5, "white")) + 
         geom_point(data = rslt, position = 'jitter',
                    aes(x = longitude, y = latitude, 
                        color = res_pls, size = abs(res_pls))) +
         scale_color_gradient2(name = "", mid = "grey", limits = range(rslt[,11:16])) +
         scale_x_continuous(limits = c(min(rslt$longitude)-2,
                                       max(rslt$longitude)+2), 
                            expand = c(0, 0)) +
         scale_y_continuous(limits = c(min(rslt$latitude)-2,
                                       max(rslt$latitude)+2), 
                            expand = c(0, 0)) +
         labs(x = "Longitude", y = "Latitude") +
         guides(size = FALSE) + ggtitle("UK-PLS")

  p1.pl <- ggmap(usMap, darken = c(.5, "white")) + 
         geom_point(data = rslt, position = 'jitter', 
                    aes(x = longitude, y = latitude, 
                        color = res_spatrf, size = abs(res_spatrf))) +
         scale_color_gradient2(name = "", mid = "grey", limits = range(rslt[,11:16])) +
         scale_x_continuous(limits = c(min(rslt$longitude)-2,
                                       max(rslt$longitude)+2), 
                            expand = c(0, 0)) +
         scale_y_continuous(limits = c(min(rslt$latitude)-2,
                                       max(rslt$latitude)+2), 
                            expand = c(0, 0)) +
    labs(x = "Longitude", y = "Latitude") +
        guides(size = FALSE) + ggtitle("SpatRF (PL)")
  
   p1.rf <- ggmap(usMap, darken = c(.5, "white")) + 
         geom_point(data = rslt, position = 'jitter', 
                    aes(x = longitude, y = latitude, 
                        color = res_rf, size = abs(res_rf))) +
         scale_color_gradient2(name = "", mid = "grey", limits = range(rslt[,11:16])) +
         scale_x_continuous(limits = c(min(rslt$longitude)-2,
                                       max(rslt$longitude)+2), 
                            expand = c(0, 0)) +
         scale_y_continuous(limits = c(min(rslt$latitude)-2,
                                       max(rslt$latitude)+2), 
                            expand = c(0, 0)) +
     labs(x = "longitude", y = "latitude") +
        guides(size = FALSE) + ggtitle("RF")
   
    p1.tprs <- ggmap(usMap, darken = c(.5, "white")) + 
         geom_point(data = rslt, position = 'jitter', 
                    aes(x = longitude, y = latitude, 
                        color = res_tprs, size = abs(res_tprs))) +
         scale_color_gradient2(name = "", mid = "grey", limits = range(rslt[,11:16])) +
         scale_x_continuous(limits = c(min(rslt$longitude)-2,
                                       max(rslt$longitude)+2), 
                            expand = c(0, 0)) +
         scale_y_continuous(limits = c(min(rslt$latitude)-2,
                                       max(rslt$latitude)+2), 
                            expand = c(0, 0)) +
      labs(x = "Longitude", y = "Latitude") +
        guides(size = FALSE) + ggtitle("TPRS")
    
    p1.rf.tprs <- ggmap(usMap) + 
         geom_point(data = rslt, position = 'jitter', 
                    aes(x = longitude, y = latitude, 
                        color = res_rf_tprs, size = abs(res_rf_tprs))) +
         scale_color_gradient2(name = "", mid = "grey", limits = range(rslt[,11:16])) +
         scale_x_continuous(limits = c(min(rslt$longitude)-2,
                                       max(rslt$longitude)+2), 
                            expand = c(0, 0)) +
         scale_y_continuous(limits = c(min(rslt$latitude)-2,
                                       max(rslt$latitude)+2), 
                            expand = c(0, 0)) +
      labs(x = "Longitude", y = "Latitude") +
        guides(size = FALSE) + ggtitle("RF-TPRS")
    
    p1.tprs.rf <- ggmap(usMap) + 
         geom_point(data = rslt, position = 'jitter', 
                    aes(x = longitude, y = latitude, 
                        color = res_tprs_rf, size = abs(res_tprs_rf))) +
         scale_color_gradient2(name = "", mid = "grey", limits = range(rslt[,11:16])) +
         scale_x_continuous(limits = c(min(rslt$longitude)-2,
                                       max(rslt$longitude)+2), 
                            expand = c(0, 0)) +
         scale_y_continuous(limits = c(min(rslt$latitude)-2,
                                       max(rslt$latitude)+2), 
                            expand = c(0, 0)) +
      labs(x = "Longitude", y = "Latitude") +
        guides(size = FALSE) + ggtitle("TPRS-RF")

  ggarrange(p1.pls, p1.pl, p1.rf, p1.tprs, p1.tprs.rf, p1.rf.tprs, 
            nrow = 2, ncol = 3, common.legend = TRUE, legend = "right") %>%
    annotate_figure(top = text_grob(tt, face = "bold", size = 16))
  # grid.arrange(arrangeGrob(p.obs,p.pls,p.pl,p.cv,nrow=2))
}

plot_res(annavg.ec, Y.pls, Y.spatrf.pl, Y.spatrf.cv, Y.rf, Y.tprs, Y.rf.tprs, Y.tprs.rf, 
         "Cross-validated Prediction Errors, EC", plltnt.ec)
# plot_res(annavg.ec, Y.pls, Y.spatrf.pl, Y.spatrf.cv, "EC residuals", plltnt.ec)
```

### OC
```{r, echo = FALSE, fig.width=24, fig.height=15}
plot_pred(annavg.oc, Y.pls.oc, Y.spatrf.pl.oc, Y.spatrf.cv.oc, "OC predictions", plltnt.ec)
```

```{r, echo = FALSE, fig.width=24, fig.height=12}
plot_res(annavg.oc, Y.pls.oc, Y.spatrf.pl.oc, Y.spatrf.cv.oc, 
         Y.rf.oc, Y.tprs.oc, Y.rf.tprs.oc, Y.tprs.rf.oc, 
         "Cross-validated Prediction Errors, OC", plltnt.ec)
```

### S
```{r, echo = FALSE, fig.width=24, fig.height=15}
plot_pred(annavg.s, Y.pls.s, Y.spatrf.pl.s, Y.spatrf.cv.s, "Cross-validated Prediction Errors, S",
          plltnt.s)
```

```{r, echo = FALSE, fig.width=24, fig.height=12}
# plot_res(annavg.s, Y.pls.s, Y.spatrf.pl.s, Y.spatrf.cv.s, "S residuals", plltnt.s)
plot_res(annavg.s, Y.pls.s, Y.spatrf.pl.s, Y.spatrf.cv.s, Y.rf.s, Y.tprs.s, Y.rf.tprs.s, Y.tprs.rf.s, 
         "Cross-validated Prediction Errors, S", plltnt.s)
```

### Si
```{r, echo = FALSE, fig.width=24, fig.height=15}
plot_pred(annavg.si, Y.pls.si, Y.spatrf.pl.si, Y.spatrf.cv.si, "Si predictions", plltnt.s)
```

```{r, echo = FALSE, fig.width=24, fig.height=12}
plot_res(annavg.si, Y.pls.si, Y.spatrf.pl.si, Y.spatrf.cv.si, 
         Y.rf.si, Y.tprs.si, Y.rf.tprs.si, Y.tprs.rf.si, 
         "Cross-validated Prediction Errors, Si", plltnt.s)
```

## Variable importance plots

### EC

```{r, fig.width=10, fig.height = 12}
library(tidyverse)

load("/Users/sicheng/Library/CloudStorage/OneDrive-UW/exposurepred/national/varimp_3_ec_full.RData")

cols = 1:ncol(X)
qs = c(.25, .5, .75)

avg.pls = avg.spatrf = matrix(NA, ncol = length(qs), nrow = ncol(X))
k = 1

for (i in cols){
  for (j in 1:length(qs)){
    avg.pls[i,j] = mean(rslt.ec$Y.pls[,k]^2, na.rm = TRUE)
    avg.spatrf[i,j] = mean(rslt.ec$Y.spatrf.cv[,k]^2, na.rm = TRUE)
    # avg.rf[i,j] = mean(rslt.pls[[k]]$Y.rf^2, na.rm = TRUE)
    # avg.rf.tprs[i,j] = mean(rslt.pls[[k]]$Y.rf.tprs^2, na.rm = TRUE)
    k = k + 1
  }
}

plotImp = function(impMat, tt, rslt = rslt.ec){
  my.alpha = .3
  # colnames(impMat) = c("Method PLS", "Method SpatRF-PL", "Method RF", "Method RF-TPRS")
  colnames(impMat) = c("Method PLS", "Method SpatRF")
  # impMat = impMat[,c(1,2)]
  
  split_cov_name <- function(dt, cov) {
    dt <- suppressWarnings(dt %>%
                             rename(cov_full_name = cov) %>%
                             mutate(
                               buffer = substr(cov_full_name, nchar(cov_full_name)-4, nchar(cov_full_name)),
                               buffer = as.numeric(ifelse(!is.na(as.integer(buffer)), buffer, NA)),
                               cov = ifelse(is.na(buffer), cov_full_name, substr(cov_full_name, 1, nchar(cov_full_name)-5) )
                             )
    ) %>%
      select(contains("cov"), buffer, everything())
    return(dt)
  }
  
  impMat = as.matrix(impMat)
  rownames(impMat) = (rslt$spatrf.varnames)
  dtNew <- impMat %>%
    as.data.frame() %>%
    rownames_to_column(var = "cov") %>%
    # rename variables if buffers
    split_cov_name(cov = "cov") %>%
      #make long format for faceting
      gather(key = "Method", value = "Importance", contains("Method")) %>%
      mutate(Method = substr(Method, 8, nchar(Method))) 
  #colnames(dtNew)[4] = "Importance"
  
  dtNew  %>%
    #buffered covariates
    drop_na(buffer) %>%
    ggplot(aes(x = Importance, y = cov)) +
    geom_point(aes(size=buffer),
               shape=1,
               alpha=my.alpha) +
    scale_size(breaks = c(min(dtNew$buffer, na.rm = T),
                          max(dtNew$buffer, na.rm = T)
    )) + #500, 5000, 10000,
    #non-buffered covariates
    geom_point(data = dtNew[is.na(dtNew$buffer),],
               alpha=my.alpha,
               aes(shape="")) +
    geom_vline(xintercept=0,
               linetype="solid",
               alpha=my.alpha) +
    facet_wrap(~ Method, scales = "free_x",
                 labeller = "label_both",
                 nrow = 1) +
    labs(y = "Geocovariate", x = "Predictor Contribution",
         shape= "non-buffer", #"proximity,\nelevation",
         #title = paste("PLS Importance:", pollutant)) +
         title = tt) +
    theme(legend.position = "bottom", text = element_text(size = 16))
}

plotImp(cbind((avg.pls[,2]) - (avg.pls[,1]), 
              (avg.spatrf[,2]) - (avg.spatrf[,1])),
        "Mean difference in prediction, 25% to 50% quantile, EC")

plotImp(cbind((avg.pls[,3]) - (avg.pls[,2]), 
              (avg.spatrf[,3]) - (avg.spatrf[,2])),
        "Mean difference in prediction, 50% to 75% quantile, EC")

plotImp(cbind((avg.pls[,3]) - (avg.pls[,1]), 
              (avg.spatrf[,3]) - (avg.spatrf[,1])),
        "Mean difference in prediction, 25% to 75% quantile, EC")

```
Largest differences between UK-PLS and Spatial RF:

```{r}
vim_25_50 = cbind((avg.pls[,2]) - (avg.pls[,1]), 
              (avg.spatrf[,2]) - (avg.spatrf[,1])) %>% #scale() %>%
  as.data.frame() %>% mutate(diff = abs(V2 - V1), 
                             varname = rslt.ec$spatrf.varnames)
vim_50_75 = cbind((avg.pls[,3]) - (avg.pls[,2]), 
              (avg.spatrf[,3]) - (avg.spatrf[,2])) %>% #scale() %>%
  as.data.frame() %>% mutate(diff = abs(V2 - V1), 
                             varname = rslt.ec$spatrf.varnames)
vim_25_75 = cbind((avg.pls[,3]) - (avg.pls[,1]), 
              (avg.spatrf[,3]) - (avg.spatrf[,1])) %>% #scale() %>%
  as.data.frame() %>% mutate(diff = abs(V2 - V1), 
                             varname = rslt.ec$spatrf.varnames)

vim_25_50 %>% top_n(diff, n = 10) %>% 
  knitr::kable(col.names = c("PLS", "SpatRF", "abs(difference)", "Variable"),
               caption = "25% to 50% quantile, EC")
vim_50_75 %>% top_n(diff, n = 10)%>% 
  knitr::kable(col.names = c("PLS", "SpatRF", "abs(difference)", "Variable"),
               caption = "50% to 75% quantile, EC")
vim_25_75 %>% top_n(diff, n = 10) %>% 
  knitr::kable(col.names = c("PLS", "SpatRF", "abs(difference)", "Variable"),
               caption = "25% to 75% quantile, EC")
```

```{r}

```

**Cleaned plots, top 10 differences**

```{r, echo = FALSE, fig.width=16, fig.height = 6}

plotImp2 = function(impMat1, impMat2, impMat3, tt, n_top = 5){
  # my.alpha = .3
  my.alpha = .8
  
  colnames(impMat1) = colnames(impMat2) = colnames(impMat3) =
    c("Method UK-PLS", "Method SpatRF")
  
  split_cov_name <- function(dt, cov) {
    dt <- suppressWarnings(dt %>%
                             rename(cov_full_name = cov) %>%
                             mutate(
                               buffer = substr(cov_full_name, nchar(cov_full_name)-4, nchar(cov_full_name)),
                               buffer = as.numeric(ifelse(!is.na(as.integer(buffer)), buffer, NA)),
                               cov = ifelse(is.na(buffer), cov_full_name, substr(cov_full_name, 1, nchar(cov_full_name)-5) )
                             )
    ) %>%
      select(contains("cov"), buffer, everything())
    
    # elevation
    dt$cov[grepl("^elev_.+_above$", dt$cov_full_name)] <- "elev_above"
    dt$cov[grepl("^elev_.+_below$", dt$cov_full_name)] <- "elev_below"
    dt$cov[grepl("^elev_.+_stdev$", dt$cov_full_name)] <- "elev_stdev"
    dt$cov[grepl("^elev_.+_at_elev$", dt$cov_full_name)] <- "elev_at_elev"
    
    dt$buffer[grepl("_1k_", dt$cov_full_name)] <- 1000
    dt$buffer[grepl("_5k_", dt$cov_full_name)] <- 5000
    
    return(dt)
  }
  
  impMat = rbind(impMat1, impMat2, impMat3) %>% as.data.frame()
  # impMat.std = rbind(impMat1.std, impMat2.std, impMat3.std) %>% as.data.frame()
  impMat$Quantiles = rep(c("25% to 50%", "50% to 75%", "25% to 75%"), 
                         each = nrow(impMat1))
  impMat$Diff = abs(impMat[,1] - impMat[,2])
  # impMat$Diff = abs(impMat.std[,1] - impMat.std[,2])
  impMat$cov = rep(colnames(X), 3)
  
  # impMat_top = impMat %>% group_by(Quantiles) %>% 
  #   slice_max(order_by = Diff, n = 5)
  
  #rownames(impMat) = rep(colnames(X), 3)
  dtNew <- impMat %>%
    as.data.frame() %>%
    #rownames_to_column(var = "cov") %>%
    # rename variables if buffers
    split_cov_name(cov = "cov") %>%
    #make long format for faceting
    gather(key = "Method", value = "Importance", contains("Method")) %>%
    mutate(Method = substr(Method, 8, nchar(Method))) 
  
  # set1 = dtNew %>% filter(Quantiles == "25% to 50%" & Method == "PLS") %>%
  #   slice_max(order_by = Diff, n = n_top) %>% select(cov) %>% c() %>% unlist()
  # set2 = dtNew %>% filter(Quantiles == "50% to 75%" & Method == "PLS") %>%
  #   slice_max(order_by = Diff, n = n_top) %>% select(cov) %>% c() %>% unlist()
  # set3 = dtNew %>% filter(Quantiles == "25% to 75%" & Method == "PLS") %>%
  #   slice_max(order_by = Diff, n = n_top) %>% select(cov) %>% c() %>% unlist()
  set1 = dtNew %>% filter(Quantiles == "25% to 50%" & Method == "UK-PLS") %>%
    slice_max(order_by = abs(Importance), n = n_top) %>% select(cov) %>% c() %>% unlist()
  set2 = dtNew %>% filter(Quantiles == "50% to 75%" & Method == "UK-PLS") %>%
    slice_max(order_by = abs(Importance), n = n_top) %>% select(cov) %>% c() %>% unlist()
  set3 = dtNew %>% filter(Quantiles == "25% to 75%" & Method == "UK-PLS") %>%
    slice_max(order_by = abs(Importance), n = n_top) %>% select(cov) %>% c() %>% unlist()
  set1.2 = dtNew %>% filter(Quantiles == "25% to 50%" & Method == "SpatRF") %>%
    slice_max(order_by = abs(Importance), n = n_top) %>% select(cov) %>% c() %>% unlist()
  set2.2 = dtNew %>% filter(Quantiles == "50% to 75%" & Method == "SpatRF") %>%
    slice_max(order_by = abs(Importance), n = n_top) %>% select(cov) %>% c() %>% unlist()
  set3.2 = dtNew %>% filter(Quantiles == "25% to 75%" & Method == "SpatRF") %>%
    slice_max(order_by = abs(Importance), n = n_top) %>% select(cov) %>% c() %>% unlist()
  # subset = set1 %>% union(set2) %>% union(set3)
  subset = set1 %>% union(set2) %>% union(set3) %>% 
    union(set1.2) %>% union(set2.2) %>% union(set3.2)
  dtNew = dtNew[dtNew$cov %in% subset, ] %>% arrange(cov)
  
  dtNew %>%
    #buffered covariates
    drop_na(buffer) %>%
    ggplot(aes(x = Importance, y = cov)) +
    geom_point(aes(size=buffer, color = Method), shape=1, alpha=my.alpha) +
    scale_size(breaks = c(min(dtNew$buffer, na.rm = T),
                          max(dtNew$buffer, na.rm = T))) + #500, 5000, 10000,
    #non-buffered covariates
    geom_point(data = dtNew[is.na(dtNew$buffer),],
               alpha=my.alpha, aes(shape="", color = Method)) +
    geom_vline(xintercept=0,
               linetype="solid",
               alpha=.3) +
    scale_y_discrete(labels=c("pop_s" = "Population", "ndvi_summer_a" = "NDVI (summer)",
                              "ndvi_q75_a" = "NDVI (annual 75%ile)",
                              "ndvi_q50_a" = "NDVI (annual median)",
                              "ndvi_winter_a" = "NDVI (winter)",
                              "m_to_ry" = "Dist to railyard",
                              "m_to_rr" = "Dist to railroad",
                              "m_to_l_airp" = "Dist to large airport",
                              "m_to_airp" = "Dist to airport",
                              "m_to_a1" = "Dist to A1 road",
                              "lu_transport_p" = "Transportation",
                              "lu_resi_p" = "Residential",
                              "lu_industrial_p" = "Industrial",
                              "lu_comm_p" = "Commercial",
                              "lu_crop_p" = "Agricultural",
                              "lu_green_p" = "Forest land",
                              "lu_oth_urban_p" = "Urban (other) land use",
                              "ll_a3_s" = "A3 road length",
                              "tl_s" = "Truck route length",
                              "rlu_evergreen_p" = "Evergreen forest",
                              "rlu_water_p" = "Water",
                              "pop10_s" = "Population",
                              "ll_a1_s" = "A1 road length",
                              "ll_a23_s" = "A2,A3 road length",
                              "rlu_dev_hi_p" = "Dev high intensity",
                              "rlu_dev_med_p" = "Dev medium intensity",
                              "log_m_to_ry" = "Log(dist) to railyard",
                              "log_m_to_l_airp" = "Log(dist) to large airport",
                              "log_m_to_airp" = "Log(dist) to airport",
                              "log_m_to_m_port" = "Log(dist) to medium port",
                              "log_m_to_a1" = "Log(dist) to A1 road",
                              "log_m_to_a1_a1_intersect" = "Log(dist) to A1/A1 intersec",
                              "log_m_to_a2_a3_intersect" = "Log(dist) to A2/A3 intersec",
                              "log_m_to_a1_a3_intersect" = "Log(dist) to A1/A3 intersec",
                              "log_m_to_truck" = "Log(dist) to truck route",
                              "log_m_to_rr" = "Log(dist) to railroad",
                              "imp_a" = "Impervious surface",
                              "elev_elevation" = "Elevation",
                              "elev_below" = "High relative elevation",
                              "elev_at_elev" = "Intermediate relative elevation",
                              "elev_above" = "Low relative elevation")) +
    facet_wrap(~ Quantiles, scales = "fixed", # scales = "free_x",
               labeller = "label_both",
               nrow = 1) +
    labs(y = "Geocovariate", x = "Predictor Contribution",
         shape= "non-buffer", #"proximity,\nelevation",
         #title = paste("PLS Importance:", pollutant)) +
         title = tt) +
    theme(legend.position = "bottom", text = element_text(size = 16))
}

plotImp2(cbind((avg.pls[,2]) - (avg.pls[,1]), 
              (avg.spatrf[,2]) - (avg.spatrf[,1])),
         cbind((avg.pls[,3]) - (avg.pls[,2]), 
              (avg.spatrf[,3]) - (avg.spatrf[,2])),
         cbind((avg.pls[,3]) - (avg.pls[,1]), 
              (avg.spatrf[,3]) - (avg.spatrf[,1])),
        "Predictors with Top Contribution, EC", n_top = 5)
```

### OC
```{r, fig.width=10, fig.height = 12}

load("/Users/sicheng/Library/CloudStorage/OneDrive-UW/exposurepred/national/varimp_3_oc.RData")

cols = 1:ncol(X)
qs = c(.25, .5, .75)

avg.pls = avg.spatrf = matrix(NA, ncol = length(qs), nrow = ncol(X))
k = 1

for (i in cols){
  for (j in 1:length(qs)){
    avg.pls[i,j] = mean(rslt.oc$Y.pls[,k]^2, na.rm = TRUE)
    avg.spatrf[i,j] = mean(rslt.oc$Y.spatrf.cv[,k]^2, na.rm = TRUE)
    # avg.rf[i,j] = mean(rslt.pls[[k]]$Y.rf^2, na.rm = TRUE)
    # avg.rf.tprs[i,j] = mean(rslt.pls[[k]]$Y.rf.tprs^2, na.rm = TRUE)
    k = k + 1
  }
}

plotImp(cbind((avg.pls[,2]) - (avg.pls[,1]), 
              (avg.spatrf[,2]) - (avg.spatrf[,1])),
        "Mean difference in prediction, 25% to 50% quantile, OC")

plotImp(cbind((avg.pls[,3]) - (avg.pls[,2]), 
              (avg.spatrf[,3]) - (avg.spatrf[,2])),
        "Mean difference in prediction, 50% to 75% quantile, OC")

plotImp(cbind((avg.pls[,3]) - (avg.pls[,1]), 
              (avg.spatrf[,3]) - (avg.spatrf[,1])),
        "Mean difference in prediction, 25% to 75% quantile, OC")
```

Largest differences between UK-PLS and Spatial RF:

```{r}
vim_25_50 = cbind((avg.pls[,2]) - (avg.pls[,1]), 
              (avg.spatrf[,2]) - (avg.spatrf[,1])) %>% #scale() %>%
  as.data.frame() %>% mutate(diff = abs(V2 - V1), 
                             varname = rslt.oc$spatrf.varnames)
vim_50_75 = cbind((avg.pls[,3]) - (avg.pls[,2]), 
              (avg.spatrf[,3]) - (avg.spatrf[,2])) %>% #scale() %>%
  as.data.frame() %>% mutate(diff = abs(V2 - V1), 
                             varname = rslt.oc$spatrf.varnames)
vim_25_75 = cbind((avg.pls[,3]) - (avg.pls[,1]), 
              (avg.spatrf[,3]) - (avg.spatrf[,1])) %>% #scale() %>%
  as.data.frame() %>% mutate(diff = abs(V2 - V1), 
                             varname = rslt.oc$spatrf.varnames)

vim_25_50 %>% top_n(diff, n = 5) %>% 
  knitr::kable(col.names = c("PLS", "SpatRF", "abs(difference)", "Variable"),
               caption = "25% to 50% quantile, OC")
vim_50_75 %>% top_n(diff, n = 5)%>% 
  knitr::kable(col.names = c("PLS", "SpatRF", "abs(difference)", "Variable"),
               caption = "50% to 75% quantile, OC")
vim_25_75 %>% top_n(diff, n = 5) %>% 
  knitr::kable(col.names = c("PLS", "SpatRF", "abs(difference)", "Variable"),
               caption = "25% to 75% quantile, OC")
```

**Cleaned plots, top 10 differences**

```{r, echo = FALSE, fig.width=16, fig.height = 6}

plotImp2(cbind((avg.pls[,2]) - (avg.pls[,1]), 
              (avg.spatrf[,2]) - (avg.spatrf[,1])),
         cbind((avg.pls[,3]) - (avg.pls[,2]), 
              (avg.spatrf[,3]) - (avg.spatrf[,2])),
         cbind((avg.pls[,3]) - (avg.pls[,1]), 
              (avg.spatrf[,3]) - (avg.spatrf[,1])),
        "Predictors with Top Contribution, OC", n_top = 5)
```

### S

```{r, fig.width=10, fig.height = 12}
load("C:/Users/Si Cheng/Dropbox (Personal)/exposurepred/national/varimp_3_s.RData")

rm(rslt.oc)

cols = 1:ncol(X)
qs = c(.25, .5, .75)

avg.pls = avg.spatrf = matrix(NA, ncol = length(qs), nrow = ncol(X))
k = 1

for (i in cols){
  for (j in 1:length(qs)){
    avg.pls[i,j] = mean(rslt.s$Y.pls[,k]^2, na.rm = TRUE)
    avg.spatrf[i,j] = mean(rslt.s$Y.spatrf.cv[,k]^2, na.rm = TRUE)
    # avg.rf[i,j] = mean(rslt.pls[[k]]$Y.rf^2, na.rm = TRUE)
    # avg.rf.tprs[i,j] = mean(rslt.pls[[k]]$Y.rf.tprs^2, na.rm = TRUE)
    k = k + 1
  }
}

plotImp(cbind((avg.pls[,2]) - (avg.pls[,1]), 
              (avg.spatrf[,2]) - (avg.spatrf[,1])),
        "Mean difference in prediction, 25% to 50% quantile, S", rslt = rslt.s)

plotImp(cbind((avg.pls[,3]) - (avg.pls[,2]), 
              (avg.spatrf[,3]) - (avg.spatrf[,2])),
        "Mean difference in prediction, 50% to 75% quantile, S", rslt = rslt.s)

plotImp(cbind((avg.pls[,3]) - (avg.pls[,1]), 
              (avg.spatrf[,3]) - (avg.spatrf[,1])),
        "Mean difference in prediction, 25% to 75% quantile, S", rslt = rslt.s)
```

Largest differences between UK-PLS and Spatial RF:

```{r}
vim_25_50 = cbind((avg.pls[,2]) - (avg.pls[,1]), 
              (avg.spatrf[,2]) - (avg.spatrf[,1])) %>% #scale() %>%
  as.data.frame() %>% mutate(diff = abs(V2 - V1), 
                             varname = rslt.s$spatrf.varnames)
vim_50_75 = cbind((avg.pls[,3]) - (avg.pls[,2]), 
              (avg.spatrf[,3]) - (avg.spatrf[,2])) %>% #scale() %>%
  as.data.frame() %>% mutate(diff = abs(V2 - V1), 
                             varname = rslt.s$spatrf.varnames)
vim_25_75 = cbind((avg.pls[,3]) - (avg.pls[,1]), 
              (avg.spatrf[,3]) - (avg.spatrf[,1])) %>% #scale() %>%
  as.data.frame() %>% mutate(diff = abs(V2 - V1), 
                             varname = rslt.s$spatrf.varnames)

vim_25_50 %>% top_n(diff, n = 5) %>% 
  knitr::kable(col.names = c("PLS", "SpatRF", "abs(difference)", "Variable"),
               caption = "25% to 50% quantile, S")
vim_50_75 %>% top_n(diff, n = 5)%>% 
  knitr::kable(col.names = c("PLS", "SpatRF", "abs(difference)", "Variable"),
               caption = "50% to 75% quantile, S")
vim_25_75 %>% top_n(diff, n = 5) %>% 
  knitr::kable(col.names = c("PLS", "SpatRF", "abs(difference)", "Variable"),
               caption = "25% to 75% quantile, S")
```

**Cleaned plots, top 10 differences**

```{r, echo = FALSE, fig.width=16, fig.height = 6}

plotImp2(cbind((avg.pls[,2]) - (avg.pls[,1]), 
              (avg.spatrf[,2]) - (avg.spatrf[,1])),
         cbind((avg.pls[,3]) - (avg.pls[,2]), 
              (avg.spatrf[,3]) - (avg.spatrf[,2])),
         cbind((avg.pls[,3]) - (avg.pls[,1]), 
              (avg.spatrf[,3]) - (avg.spatrf[,1])),
        "Predictors with Top Contribution, S", n_top = 5)
```

### Si

```{r, fig.width=10, fig.height = 12}
load("/Users/sicheng/Library/CloudStorage/OneDrive-UW/exposurepred/national/varimp_si.RData")

cols = 1:ncol(X)
qs = c(.25, .5, .75)

avg.pls = avg.spatrf = matrix(NA, ncol = length(qs), nrow = ncol(X))
k = 1

for (i in cols){
  for (j in 1:length(qs)){
    avg.pls[i,j] = mean(rslt.si$Y.pls[,k]^2, na.rm = TRUE)
    avg.spatrf[i,j] = mean(rslt.si$Y.spatrf.cv[,k]^2, na.rm = TRUE)
    # avg.rf[i,j] = mean(rslt.pls[[k]]$Y.rf^2, na.rm = TRUE)
    # avg.rf.tprs[i,j] = mean(rslt.pls[[k]]$Y.rf.tprs^2, na.rm = TRUE)
    k = k + 1
  }
}

plotImp(cbind((avg.pls[,2]) - (avg.pls[,1]), 
              (avg.spatrf[,2]) - (avg.spatrf[,1])),
        "Mean difference in prediction, 25% to 50% quantile, Si", rslt = rslt.si)

plotImp(cbind((avg.pls[,3]) - (avg.pls[,2]), 
              (avg.spatrf[,3]) - (avg.spatrf[,2])),
        "Mean difference in prediction, 50% to 75% quantile, Si", rslt = rslt.si)

plotImp(cbind((avg.pls[,3]) - (avg.pls[,1]), 
              (avg.spatrf[,3]) - (avg.spatrf[,1])),
        "Mean difference in prediction, 25% to 75% quantile, Si", rslt = rslt.si)
```

Largest differences between UK-PLS and Spatial RF:

```{r}
vim_25_50 = cbind((avg.pls[,2]) - (avg.pls[,1]), 
              (avg.spatrf[,2]) - (avg.spatrf[,1])) %>% #scale() %>%
  as.data.frame() %>% mutate(diff = abs(V2 - V1), 
                             varname = rslt.s$spatrf.varnames)
vim_50_75 = cbind((avg.pls[,3]) - (avg.pls[,2]), 
              (avg.spatrf[,3]) - (avg.spatrf[,2])) %>% #scale() %>%
  as.data.frame() %>% mutate(diff = abs(V2 - V1), 
                             varname = rslt.s$spatrf.varnames)
vim_25_75 = cbind((avg.pls[,3]) - (avg.pls[,1]), 
              (avg.spatrf[,3]) - (avg.spatrf[,1])) %>% #scale() %>%
  as.data.frame() %>% mutate(diff = abs(V2 - V1), 
                             varname = rslt.s$spatrf.varnames)

vim_25_50 %>% top_n(diff, n = 5) %>% 
  knitr::kable(col.names = c("PLS", "SpatRF", "abs(difference)", "Variable"),
               caption = "25% to 50% quantile, Si")
vim_50_75 %>% top_n(diff, n = 5)%>% 
  knitr::kable(col.names = c("PLS", "SpatRF", "abs(difference)", "Variable"),
               caption = "50% to 75% quantile, Si")
vim_25_75 %>% top_n(diff, n = 5) %>% 
  knitr::kable(col.names = c("PLS", "SpatRF", "abs(difference)", "Variable"),
               caption = "25% to 75% quantile, Si")
```

**Cleaned plots, top 10 differences**

```{r, echo = FALSE, fig.width=16, fig.height = 6}

plotImp2(cbind((avg.pls[,2]) - (avg.pls[,1]), 
              (avg.spatrf[,2]) - (avg.spatrf[,1])),
         cbind((avg.pls[,3]) - (avg.pls[,2]), 
              (avg.spatrf[,3]) - (avg.spatrf[,2])),
         cbind((avg.pls[,3]) - (avg.pls[,1]), 
              (avg.spatrf[,3]) - (avg.spatrf[,1])),
        "Predictors with Top Contribution, Si", n_top = 10)
```
